### 2021-04-26 R/exploration.Rmd
init data exploration

---
title: "JStudy exploratory data analysis"
author: "Monica Thieu"
date: "4/25/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(tidyverse)
require(magrittr)
j_recall <- read_csv(here::here("ignore", "data", "task_jeopardy_recall.csv"))
j_meta <- read_csv(here::here("ignore", "data", "task_jeopardy_meta_states.csv"))
encoding <- read_csv(here::here("ignore", "data", "task_encoding.csv"))
r_facts <- read_csv(here::here("ignore", "data", "task_retrieval_facts.csv"))
r_pics <- read_csv(here::here("ignore", "data", "task_retrieval_pics.csv"))
r_source <- read_csv(here::here("ignore", "data", "task_retrieval_source.csv"))
q_demos  <- read_csv(here::here("ignore", "data", "q_trivia_demos.csv"))
q_posttask  <- read_csv(here::here("ignore", "data", "q_posttask.csv"))
```

## Trivia expertise

```{r}
j_recall %>% 
  group_by(subj_num) %>% 
  summarize(score = mean(acc_recall))
```

```{r}
j_recall %>% 
  mutate(acc_recall = as.integer(acc_recall),
         resp_conf = (resp_conf - 50) / 100) %>% 
  glm(acc_recall ~ resp_conf, family = "binomial", data = .) %>% 
  summary()
```

```{r}
j_recall %>% 
  group_by(subj_num) %>% 
  ggplot(aes(x = resp_conf, y = as.integer(acc_recall))) + 
  geom_point() + 
  geom_smooth(method = "glm", method.args = list(family = "binomial")) + 
  facet_wrap(~subj_num)
```

## Metamemory states

```{r}
j_meta %>%
  left_join(j_recall %>% select(subj_num, answer, acc_recall)) %>% 
  filter(acc_recall) %>% 
  ggplot(aes(x = factor(subj_num), fill = resp)) +
  geom_bar()
```

```{r}
j_meta %>%
  left_join(j_recall %>% select(subj_num, answer, acc_recall)) %>% 
  filter(acc_recall) %>% 
  count(subj_num, resp) %>% 
  group_by(subj_num) %>% 
  mutate(prob = n / sum(n)) %>% 
  ggplot(aes(x = factor(subj_num), y = prob, fill = resp)) +
  geom_col()
```
```{r}
j_meta %>%
  left_join(j_recall %>% select(subj_num, answer, acc_recall)) %>% 
  filter(!acc_recall) %>% 
  ggplot(aes(x = factor(subj_num), fill = resp)) +
  geom_bar()
```

### Museum task retrieval

```{r}
r_pics %>%
  ggplot(aes(x = resp, fill = factor(is_old))) +
  geom_histogram(position = "identity", alpha = 0.5, bins = 20) + facet_wrap(~ subj_num)
```

```{r}
r_source %>%
  ggplot(aes(x = resp, fill = factor(encoding_block))) +
  geom_histogram(position = "identity", alpha = 0.5, bins = 20) + facet_wrap(~ subj_num)
```

### Source helpage

```{r}
retrieval <- r_facts %>%
  select(subj_num, group, group_trial_num, retrieval_block, retrieval_trial_num, category, acc_recall, rt_recall = rt) %>% 
  left_join(r_pics %>%
              mutate(resp = resp - 50,
                     acc = if_else(is_old == resp_binary, abs(resp), -abs(resp))) %>%
              select(subj_num, group, group_trial_num, category, pic_is_old = is_old, acc_pic = acc, rt_pic = rt)) %>% 
  left_join(r_source %>%
              mutate(resp = resp - 50,
                     acc = if_else(encoding_block == resp_binary, abs(resp), -abs(resp))) %>%
              select(subj_num, group, group_trial_num, category, acc_source = acc, rt_source = rt)) %>%
  left_join(encoding %>% 
              select(subj_num, group, group_trial_num, category, interest = resp)) %>% 
  mutate(retrieval_block = case_when(retrieval_block <= 4 ~ "ep_same_sem_same",
                                     retrieval_block <= 8 ~ "ep_same_sem_diff",
                                     retrieval_block <= 12 ~ "ep_diff_sem_same",
                                     retrieval_block <= 16 ~ "ep_diff_sem_diff",
                                     TRUE ~ NA_character_))
```

```{r}
retrieval %>% 
  mutate(acc_recall = if_else(acc_recall == 0.5, 1, acc_recall)) %>%
  ggplot(aes(x = interest, y = acc_recall)) + 
  geom_point() + 
  geom_smooth(method = "glm", method.args = list(family = "binomial")) + 
  facet_wrap(~subj_num)
```

```{r}
retrieval %>% 
  mutate(acc_recall = if_else(acc_recall == 0.5, 1, acc_recall)) %>%
  ggplot(aes(x = acc_pic, y = acc_recall)) + 
  geom_point() + 
  geom_smooth(method = "glm", method.args = list(family = "binomial")) + 
  facet_wrap(~subj_num)
```

```{r}
retrieval %>% 
  mutate(acc_recall = if_else(acc_recall == 0.5, 1, acc_recall)) %>%
  ggplot(aes(x = acc_source, y = acc_recall)) + 
  geom_point() + 
  geom_smooth(method = "glm", method.args = list(family = "binomial")) + 
  facet_wrap(~subj_num)
```

```{r}
retrieval %>% 
  ggplot(aes(x = interest, y = acc_source)) + 
  geom_point() + 
  geom_smooth(method = "lm") + 
  facet_wrap(~subj_num)
```

```{r}
retrieval %>% 
  ggplot(aes(x = acc_source, y = acc_pic)) + 
  geom_point() + 
  geom_smooth(method = "lm") + 
  facet_wrap(~subj_num)
```