### 2021-05-07 R/exploration.Rmd
make it into xaringan and nice ish

---
title: "JStudy exploratory data analysis"
subtitle: "with xaringan"
author: "Monica Thieu & Lauren Wilkins"
institute: "Columbia University Dept of Psychology"
date: "`r Sys.Date()`"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, dev = "svglite", message = FALSE, warning = FALSE, fig.asp = 3/4)
require(tidyverse)
require(magrittr)
j_recall <- read_csv(here::here("ignore", "data", "task_jeopardy_recall.csv"))

j_exp <- j_recall %>% 
  group_by(subj_num) %>% 
  summarize(j_score = mean(acc_recall))

j_recall %<>%
  left_join(j_exp, by = "subj_num") %>% 
  mutate(subj_num = fct_reorder(as.character(subj_num), j_score))

j_meta <- read_csv(here::here("ignore", "data", "task_jeopardy_meta_states.csv")) %>% 
  left_join(j_exp, by = "subj_num") %>% 
  mutate(subj_num = fct_reorder(as.character(subj_num), j_score))
encoding <- read_csv(here::here("ignore", "data", "task_encoding.csv")) %>% 
  left_join(j_exp, by = "subj_num") %>% 
  mutate(subj_num = fct_reorder(as.character(subj_num), j_score))
r_facts <- read_csv(here::here("ignore", "data", "task_retrieval_facts.csv")) %>% 
  left_join(j_exp, by = "subj_num") %>% 
  mutate(subj_num = fct_reorder(as.character(subj_num), j_score))
r_pics_oldnew <- read_csv(here::here("ignore", "data", "task_retrieval_pics_oldnew.csv")) %>% 
  left_join(j_exp, by = "subj_num") %>% 
  mutate(subj_num = fct_reorder(as.character(subj_num), j_score))
r_pics_2afc <- read_csv(here::here("ignore", "data", "task_retrieval_pics.csv")) %>% 
  left_join(j_exp, by = "subj_num") %>% 
  mutate(subj_num = fct_reorder(as.character(subj_num), j_score))
r_source <- read_csv(here::here("ignore", "data", "task_retrieval_source.csv")) %>% 
  left_join(j_exp, by = "subj_num") %>% 
  mutate(subj_num = fct_reorder(as.character(subj_num), j_score))
q_demos  <- read_csv(here::here("ignore", "data", "q_trivia_demos.csv")) %>% 
  left_join(j_exp, by = "subj_num") %>% 
  mutate(subj_num = fct_reorder(as.character(subj_num), j_score))
q_posttask  <- read_csv(here::here("ignore", "data", "q_posttask.csv"))
```

class: inverse, middle

# Trivia expertise assessment

---

```{r}
q_demos %>% 
  ggplot(aes(x = trivia_skill_self, y = j_score)) +
  geom_point(size = 3) +
  labs(x = "Self-reported trivia expertise (relative to 'average person')",
       y = "Expertise score on Jeopardy audition test") +
  theme_bw()
```

---

## Trivia recall by confidence

```{r}
j_recall %>% 
  ggplot(aes(x = resp_conf, y = as.integer(acc_recall), color = j_score, fill = j_score)) + 
  geom_jitter(alpha = 0.3, height = 0.05) + 
  geom_smooth(aes(group = subj_num), method = "glm", method.args = list(family = "binomial"), alpha = 0.05) + 
  scale_color_viridis_c() +
  scale_fill_viridis_c() +
  labs(x = "Recall confidence", y = "Recall accuracy (strict)", color = "Expertise score", fill = "Expertise score") +
  theme_bw()
```

---

## Metamemory state counts

```{r}
j_meta %>%
  left_join(j_recall %>% select(subj_num, answer, acc_recall)) %>% 
  mutate(acc_recall = recode(as.character(acc_recall),
                             `FALSE` = "incorrect",
                             `TRUE` = "correct")) %>% 
  ggplot(aes(x = subj_num, fill = resp)) +
  geom_bar() +
  scale_x_discrete(guide = guide_axis(angle = 30)) +
  facet_grid(~acc_recall) +
  labs(x = "Participant", y = "# facts labeled with that memory state", fill = "Memory state") +
  theme_bw()
```

---

## Metamemory state relative percentages

```{r}
j_meta %>%
  left_join(j_recall %>% select(subj_num, answer, acc_recall)) %>% 
  mutate(acc_recall = recode(as.character(acc_recall),
                             `FALSE` = "incorrect",
                             `TRUE` = "correct")) %>% 
  count(subj_num, acc_recall, resp) %>% 
  group_by(subj_num, acc_recall) %>% 
  mutate(prob = n / sum(n)) %>% 
  ggplot(aes(x = subj_num, y = prob, fill = resp)) +
  geom_col() +
  scale_x_discrete(guide = guide_axis(angle = 30)) +
  facet_grid(~acc_recall) +
  labs(x = "Participant", y = "Percent facts labeled with that memory state", fill = "Memory state") +
  theme_bw()
```

---

class: inverse, middle

# Museum task retrieval

---

## Fact recall as a function of question order/chunk type

```{r}
r_facts %>% 
  ggplot(aes(x = retrieval_trial_num, y = acc_recall)) +
  geom_rect(aes(x = NULL, y = NULL, xmin = xmin, xmax = xmax, fill = interaction(ep_is_same, sem_is_same)),
            data = r_facts %>%
              mutate(ep_is_same = recode(ep_is_same,
                                         `0` = "ep diff",
                                         `1` = "ep same"),
                     sem_is_same = recode(sem_is_same,
                                          `0` = "sem diff",
                                          `1` = "sem same")) %>% 
              select(subj_num, retrieval_block, retrieval_trial_num, ep_is_same, sem_is_same) %>%
              group_by(subj_num, retrieval_block) %>%
              filter(retrieval_trial_num %in% range(retrieval_trial_num)) %>%
              mutate(xtype = if_else(retrieval_trial_num == min(retrieval_trial_num), "xmin", "xmax")) %>%
              ungroup() %>%
              pivot_wider(names_from = xtype, values_from = retrieval_trial_num) %>% 
              mutate(xmin = xmin - 0.5,
                     xmax = xmax + 0.5),
            ymin = 0, ymax = 1, alpha = 0.5) +
  geom_line() +
  geom_point() +
  scale_colour_viridis_b() +
  scale_y_continuous(n.breaks = 3) +
  facet_grid(subj_num ~ .) +
  labs(x = "Fact recall trial (beginning to end)", y = "Fact recall accuracy", fill = "Chunk type") +
  theme_bw()
```

---

## Fact recall by chunk type

```{r}
r_facts %>% 
  mutate(ep_is_same = recode(ep_is_same,
                                         `0` = "ep diff",
                                         `1` = "ep same"),
                     sem_is_same = recode(sem_is_same,
                                          `0` = "sem diff",
                                          `1` = "sem same")) %>% 
  group_by(subj_num, j_score, ep_is_same, sem_is_same) %>% 
  summarize(mean_acc = mean(acc_recall)) %>% 
  ggplot(aes(x = interaction(ep_is_same, sem_is_same), y = mean_acc,
             group = subj_num,
             color = j_score)) +
  geom_line() +
  geom_point() +
  scale_x_discrete(guide = guide_axis(angle = 30)) +
  scale_color_viridis_c() +
  labs(x = "Chunk type", y = "Percent facts correctly recalled", color = "Expertise score") +
  theme_bw()
```

---

## Photo recognition

Old setup: A-B' old-new recognition with similar lures as foils

```{r}
r_pics_oldnew %>%
  ggplot(aes(x = factor(subj_num), y = resp, fill = factor(is_old))) +
  geom_hline(yintercept = 0, linetype = "dotted") +
  geom_boxplot() +
  labs(x = "Participant", y = "Accuracy with confidence (positive = correct)") +
  theme_bw()
```

```{r}
r_pics_oldnew %>%
  mutate(say_old = case_when(is_old == 0 & resp_binary == 1 ~ 0L,
                             is_old == 0 & resp_binary == 0 ~ 1L,
                             is_old == 1 & resp_binary == 0 ~ 0L,
                             is_old == 1 & resp_binary == 1 ~ 1L,
                             TRUE ~ NA_integer_)) %>% 
  nest(data = -subj_num) %>% 
  mutate(coefs = map(data, ~glm(say_old ~ is_old, family = binomial(link = "logit"), data = .) %>% 
                       broom::tidy())) %>% 
  select(subj_num, coefs) %>% 
  unnest(coefs) %>% 
  filter(term == "is_old")
```

---

### Current setup: A-A' 2AFC with similar lure

```{r}
r_pics_2afc %>%
  ggplot(aes(x = factor(subj_num), y = resp, color = j_score)) +
  geom_hline(yintercept = 0, linetype = "dotted") +
  geom_boxplot() +
  scale_colour_viridis_c() +
  guides(color = FALSE) +
  labs(x = "Participant", y = "Accuracy with confidence (positive = correct)") +
  theme_bw()
```

---

### Source memory for which museum

```{r}
r_source %>%
  ggplot(aes(x = factor(subj_num), y = resp, color = j_score)) +
  geom_hline(yintercept = 0, linetype = "dotted") +
  geom_boxplot() +
  scale_colour_viridis_c() +
  guides(color = FALSE) +
  labs(x = "Participant", y = "Accuracy with confidence (positive = correct)") +
  theme_bw()
```

---

### Source helpage

```{r}
retrieval <- r_facts %>%
  select(subj_num, j_score, group, group_trial_num, retrieval_block, ep_is_same, sem_is_same, retrieval_trial_num, category, acc_recall, rt_start_recall = rt_start, rt_end_recall = rt_end) %>% 
  left_join(r_pics_oldnew %>%
              select(subj_num, group, group_trial_num, category, pic_is_old = is_old, resp_pic = resp, rt_pic = rt) %>% 
              bind_rows(r_pics_2afc %>% 
                          select(subj_num, group, group_trial_num, category, resp_pic = resp, rt_pic = rt))) %>% 
  left_join(r_source %>%
              select(subj_num, group, group_trial_num, category, resp_source = resp, rt_source = rt)) %>%
  left_join(encoding %>% 
              select(subj_num, group, group_trial_num, category, interest = resp)) %>% 
  mutate(retrieval_block_type = case_when(retrieval_block <= 4 ~ "ep_same_sem_same",
                                          retrieval_block <= 8 ~ "ep_same_sem_diff",
                                          retrieval_block <= 12 ~ "ep_diff_sem_same",
                                          retrieval_block <= 16 ~ "ep_diff_sem_diff",
                                          TRUE ~ NA_character_))
```

---

### Does successful photo retrieval predict successful fact retrieval?

```{r}
retrieval %>% 
  mutate(acc_recall = if_else(acc_recall == 0.5, 1, acc_recall)) %>%
  ggplot(aes(x = resp_pic, y = acc_recall, color = j_score, fill = j_score)) + 
  geom_jitter(alpha = 0.3, height = 0.05) + 
  geom_smooth(aes(group = subj_num), method = "glm", method.args = list(family = "binomial"), alpha = 0.05) + 
  scale_color_viridis_c() +
  scale_fill_viridis_c() +
  labs(x = "Photo recognition accuracy", y = "Fact recall accuracy (generous)") +
  theme_bw()
```

---

### Does successful source retrieval predict successful fact retrieval?

```{r}
retrieval %>% 
  mutate(acc_recall = if_else(acc_recall == 0.5, 1, acc_recall)) %>%
  ggplot(aes(x = resp_source, y = acc_recall, color = j_score, fill = j_score)) + 
  geom_jitter(alpha = 0.3, height = 0.05) + 
  geom_smooth(aes(group = subj_num), method = "glm", method.args = list(family = "binomial"), alpha = 0.05) + 
  scale_color_viridis_c() +
  scale_fill_viridis_c() +
  labs(x = "Source recognition accuracy", y = "Fact recall accuracy (generous)") +
  theme_bw()
```

---

### Does interest at encoding predict successful _fact_ retrieval?

```{r}
retrieval %>% 
  mutate(acc_recall = if_else(acc_recall == 0.5, 1, acc_recall)) %>%
  ggplot(aes(x = interest, y = acc_recall, color = j_score, fill = j_score)) + 
  geom_jitter(alpha = 0.3, height = 0.05) + 
  geom_smooth(aes(group = subj_num), method = "glm", method.args = list(family = "binomial"), alpha = 0.05) + 
  scale_color_viridis_c() +
  scale_fill_viridis_c() +
  labs(x = "Interest in fact at encoding", y = "Fact recall accuracy (generous)") +
  theme_bw()
```

---

### Does interest at encoding predict successful _photo_ retrieval?

```{r}
retrieval %>% 
  ggplot(aes(x = interest, y = resp_pic, color = j_score, fill = j_score)) + 
  geom_point() + 
  geom_smooth(method = "lm") + 
  facet_wrap(~subj_num) +
  scale_color_viridis_c() +
  scale_fill_viridis_c() +
  labs(x = "Interest in fact at encoding", y = "Photo recognition accuracy") +
  theme_bw()
```

---

### Does interest at encoding predict successful _source_ retrieval?

```{r}
retrieval %>% 
  ggplot(aes(x = interest, y = resp_source, color = j_score, fill = j_score)) + 
  geom_point() + 
  geom_smooth(method = "lm") + 
  facet_wrap(~subj_num) +
  scale_color_viridis_c() +
  scale_fill_viridis_c() +
  labs(x = "Interest in fact at encoding", y = "Source recognition accuracy") +
  theme_bw()
```

---

### Does successful _source_ retrieval predict successful _photo_ retrieval?

```{r}
retrieval %>% 
  ggplot(aes(x = resp_source, y = resp_pic, color = j_score, fill = j_score)) + 
  geom_point() + 
  geom_smooth(method = "lm") + 
  facet_wrap(~subj_num) +
  scale_color_viridis_c() +
  scale_fill_viridis_c() +
  labs(x = "Source recognition accuracy", y = "Photo recognition accuracy") +
  theme_bw()
```

---

```{r}
retrieval %>% 
  ggplot(aes(x = log(rt_pic), y = resp_pic)) + 
  geom_point() + 
  geom_smooth(method = "lm") + 
  facet_wrap(~subj_num, scales = "free_x")
```

```{r}
retrieval %>% 
  ggplot(aes(x = log(rt_source), y = resp_source)) + 
  geom_point() + 
  geom_smooth(method = "lm") + 
  facet_wrap(~subj_num, scales = "free_x")
```