### 2021-06-30 R/study_sims.Rmd
move file, reorganize large bootstrap

---
title: "scratch simulations"
author: "Monica Thieu"
date: "4/9/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(tidyverse)
require(magrittr)
j_recall <- read_csv(here::here("ignore", "data", "task_jeopardy_recall.csv"))

j_exp <- j_recall %>% 
  group_by(subj_num) %>% 
  summarize(j_score = mean(acc_recall))

j_recall %<>%
  left_join(j_exp, by = "subj_num") %>% 
  mutate(subj_num = fct_reorder(as.character(subj_num), j_score))

j_meta <- read_csv(here::here("ignore", "data", "task_jeopardy_meta_states.csv")) %>% 
  left_join(j_exp, by = "subj_num") %>% 
  mutate(subj_num = fct_reorder(as.character(subj_num), j_score))
j_meta_desc <- read_csv(here::here("ignore", "data", "task_jeopardy_meta_descriptions.csv")) %>% 
  left_join(j_exp, by = "subj_num") %>% 
  mutate(subj_num = fct_reorder(as.character(subj_num), j_score))
encoding <- read_csv(here::here("ignore", "data", "task_encoding.csv")) %>% 
  left_join(j_exp, by = "subj_num") %>% 
  mutate(subj_num = fct_reorder(as.character(subj_num), j_score))
r_facts <- read_csv(here::here("ignore", "data", "task_retrieval_facts.csv")) %>% 
  left_join(j_exp, by = "subj_num") %>% 
  mutate(subj_num = fct_reorder(as.character(subj_num), j_score))
r_pics <- read_csv(here::here("ignore", "data", "task_retrieval_pics.csv")) %>% 
  left_join(j_exp, by = "subj_num") %>% 
  mutate(subj_num = fct_reorder(as.character(subj_num), j_score))
r_source <- read_csv(here::here("ignore", "data", "task_retrieval_source.csv")) %>% 
  left_join(j_exp, by = "subj_num") %>% 
  mutate(subj_num = fct_reorder(as.character(subj_num), j_score))
q_demos  <- read_csv(here::here("ignore", "data", "q_trivia_demos.csv")) %>% 
  left_join(j_exp, by = "subj_num") %>% 
  mutate(subj_num = fct_reorder(as.character(subj_num), j_score))

retrieval <- r_facts %>%
  select(subj_num, j_score, group, group_trial_num, retrieval_block, ep_is_same, sem_is_same, retrieval_trial_num, category, acc_recall, rt_start_recall = rt_start, rt_end_recall = rt_end) %>% 
  left_join(r_pics %>% 
              select(subj_num, group, group_trial_num, category, resp_pic = resp, rt_pic = rt)) %>% 
  left_join(r_source %>%
              select(subj_num, group, group_trial_num, category, resp_source = resp, rt_source = rt)) %>%
  left_join(encoding %>% 
              select(subj_num, group, group_trial_num, category, interest = resp)) %>% 
  mutate(retrieval_block_type = case_when(retrieval_block <= 4 ~ "ep_same_sem_same",
                                          retrieval_block <= 8 ~ "ep_same_sem_diff",
                                          retrieval_block <= 12 ~ "ep_diff_sem_same",
                                          retrieval_block <= 16 ~ "ep_diff_sem_diff",
                                          TRUE ~ NA_character_))
```

```{r}
stuff <- tibble(id = 1:20) %>%
  mutate(if_else(id %% 2 == 0, "trivia", "control"))
```

# Just Jeopardy pre-test expertise

## Parameter recovery within subject

```{r}
j_recall %>% 
  ggplot(aes(x = resp_conf)) +
  geom_histogram() +
  facet_wrap(~subj_num)
```

```{r}
j_recall %>% 
  mutate(acc_recall = as.integer(acc_recall),
         resp_conf = resp_conf / 100) %>% 
  nest(data = -c(subj_num, j_score)) %>% 
  mutate(coefs = map(data, ~glm(acc_recall ~ resp_conf, family = binomial(link = "logit"), data = .) %>% 
                       broom::tidy()
                       )) %>% 
  select(-data) %>% 
  unnest(coefs) %>% 
  select(-statistic, -p.value) %>% 
  mutate(term = recode(term, `(Intercept)` = "intercept"))
```

# Museum task

## Interest ratings

### INTERACTION: Does retrieval ~ encoding interest depend on expertise?

```{r}
retrieval %>% 
  mutate(acc_recall = if_else(acc_recall == 0.5, 1, acc_recall)) %>%
  ggplot(aes(x = interest, y = acc_recall, color = j_score, fill = j_score)) + 
  geom_jitter(alpha = 0.3, height = 0.05) + 
  geom_smooth(aes(group = subj_num), method = "glm", method.args = list(family = "binomial"), alpha = 0.05) + 
  scale_color_viridis_c() +
  scale_fill_viridis_c() +
  labs(x = "Interest in fact at encoding", y = "Fact recall accuracy (generous)") +
  theme_bw()
```

```{r}
retrieval %>% 
  mutate(acc_recall = if_else(acc_recall == 0.5, 1, acc_recall),
         interest = (interest - 50) / 100) %>% 
  nest(data = -c(subj_num, j_score)) %>% 
  mutate(model = map(data, ~glm(acc_recall ~ interest, family = binomial(link = "logit"), data = .)),
         coefs = map(model, broom::tidy)) %>% 
  select(subj_num, j_score, coefs) %>% 
  unnest() %>% 
  select(-statistic, -p.value) %>% 
  mutate(term = recode(term, `(Intercept)` = "intercept")) %>% 
  pivot_wider(names_from = term, values_from = c(estimate, std.error)) %>% 
  ggplot(aes(x = estimate_interest, y = std.error_interest, color = j_score)) +
  geom_point() +
  scale_colour_viridis_c()
```

```{r}
retrieval %>% 
  mutate(acc_recall = if_else(acc_recall == 0.5, 1, acc_recall),
         interest = (interest - 50) / 100) %>% 
  nest(data = -c(subj_num, j_score)) %>% 
  mutate(model = map(data, ~glm(acc_recall ~ interest, family = binomial(link = "logit"), data = .)),
         coefs = map(model, broom::tidy)) %>% 
  select(subj_num, j_score, coefs) %>% 
  unnest() %>% 
  select(-statistic, -p.value) %>% 
  mutate(term = recode(term, `(Intercept)` = "intercept")) %>% 
  pivot_wider(names_from = term, values_from = c(estimate, std.error)) %>% 
  ggplot(aes(x = estimate_interest, y = std.error_interest)) +
  geom_point()
```

P(correct recall for given subject) ~ invlogit(intercept fixed + intercept random + interest + j)

```{r}
glmer_retrieval_by_interest_expertise <- retrieval %>% 
  mutate(acc_recall = if_else(acc_recall == 0.5, 1, acc_recall),
         interest = (interest - 50) / 100,
         j_score = j_score - 0.5) %>% 
  lme4::glmer(acc_recall ~ interest * j_score + (1 | subj_num),
              family = binomial(link = "logit"),
              data = .)

coefs_retrieval_by_interest_expertise <- glmer_retrieval_by_interest_expertise %>%
  broom.mixed::tidy() %>%
  select(effect:estimate)
```

```{r}

```

###