### 2021-07-05 slides/data_checking.Rmd

---
title: "Exploratory data analysis"
subtitle: "J-Study"
author: "Monica Thieu & Lauren Wilkins"
institute: "Columbia University Dept of Psychology"
date: "`r Sys.Date()`"
output:
  xaringan::moon_reader:
    css: ["theme.css", "fonts.css"]
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, dev = "svglite", message = FALSE, warning = FALSE, fig.asp = 3/4)
require(tidyverse)
require(magrittr)

demos_ids <- read_csv(here::here("ignore", "data", "demos_ids.csv"))
q_google_demos <- read_csv(here::here("ignore", "data", "q_google_demos.csv")) %>% 
  right_join(demos_ids, by = "demographics_id")

j_recall <- read_csv(here::here("ignore", "data", "task_jeopardy_recall.csv")) %>% 
  mutate(resp_conf = if_else(is.na(resp_recall), NA_real_, resp_conf))

j_exp <- j_recall %>% 
  group_by(subj_num) %>% 
  summarize(j_score = mean(acc_recall))

j_recall %<>%
  left_join(j_exp, by = "subj_num") %>% 
  mutate(subj_num = fct_reorder(as.character(subj_num), j_score))

j_meta <- read_csv(here::here("ignore", "data", "task_jeopardy_meta_states.csv")) %>% 
  left_join(j_exp, by = "subj_num") %>% 
  mutate(subj_num = fct_reorder(as.character(subj_num), j_score),
         resp = fct_relevel(resp, "Learning Memory", after = Inf))
j_meta_desc <- read_csv(here::here("ignore", "data", "task_jeopardy_meta_descriptions.csv")) %>% 
  left_join(j_exp, by = "subj_num") %>% 
  mutate(subj_num = fct_reorder(as.character(subj_num), j_score))
encoding <- read_csv(here::here("ignore", "data", "task_encoding.csv")) %>% 
  left_join(j_exp, by = "subj_num") %>% 
  mutate(subj_num = fct_reorder(as.character(subj_num), j_score),
         already_knew = fct_recode(already_knew,
                               none = "Did not know this before",
                               some = "Knew some of this before",
                               all = "Knew all of this before"))
r_facts <- read_csv(here::here("ignore", "data",  "task_retrieval_facts.csv")) %>% 
  left_join(j_exp, by = "subj_num") %>% 
  mutate(subj_num = fct_reorder(as.character(subj_num), j_score))
r_pics <- read_csv(here::here("ignore", "data", "task_retrieval_pics.csv")) %>% 
  left_join(j_exp, by = "subj_num") %>% 
  mutate(subj_num = fct_reorder(as.character(subj_num), j_score))
r_source <- read_csv(here::here("ignore", "data", "task_retrieval_source.csv")) %>% 
  left_join(j_exp, by = "subj_num") %>% 
  mutate(subj_num = fct_reorder(as.character(subj_num), j_score))
q_trivia_demos  <- read_csv(here::here("ignore", "data", "q_trivia_demos.csv")) %>% 
  left_join(j_exp, by = "subj_num") %>% 
  mutate(subj_num = fct_reorder(as.character(subj_num), j_score))

q_posttask  <- read_csv(here::here("ignore", "data", "q_posttask.csv"))

retrieval <- r_facts %>%
  select(subj_num, j_score, group, group_trial_num, retrieval_block, ep_is_same, sem_is_same, retrieval_trial_num, category, acc_recall, rt_start_recall = rt_start, rt_end_recall = rt_end) %>% 
  left_join(r_pics %>% 
              select(subj_num, group, group_trial_num, category, resp_pic = resp, rt_pic = rt)) %>% 
  left_join(r_source %>%
              select(subj_num, group, group_trial_num, category, resp_source = resp, rt_source = rt)) %>%
  left_join(encoding %>% 
              select(subj_num, group, group_trial_num, category, interest = resp, already_knew)) %>% 
  mutate(retrieval_block_type = case_when(retrieval_block <= 4 ~ "ep_same_sem_same",
                                          retrieval_block <= 8 ~ "ep_same_sem_diff",
                                          retrieval_block <= 12 ~ "ep_diff_sem_same",
                                          retrieval_block <= 16 ~ "ep_diff_sem_diff",
                                          TRUE ~ NA_character_),
         encoding_trial_num = if_else(group == "academic",
                                      group_trial_num * 2 - 1,
                                      group_trial_num * 2))
```

class: inverse, middle

# Trivia expertise assessment

---

### Do Ps accurately assess their own trivia skill?

```{r}
q_demos %>% 
  ggplot(aes(x = trivia_skill_self, y = j_score)) +
  geom_abline(slope = .01, intercept = 0, linetype = "dotted") +
  geom_point(size = 3) +
  expand_limits(y = 0) +
  labs(x = "Self-reported trivia expertise (relative to 'average person')",
       y = "Expertise score on Jeopardy audition test") +
  theme_bw()
```

---

### Trivia recall by confidence

```{r}
j_recall %>% 
  ggplot(aes(x = resp_conf, y = as.integer(acc_recall), color = j_score, fill = j_score)) + 
  geom_jitter(alpha = 0.3, height = 0.05) + 
  geom_smooth(aes(group = subj_num), method = "glm", method.args = list(family = "binomial"), se = FALSE) + 
  scale_color_viridis_c() +
  scale_fill_viridis_c() +
  labs(x = "Recall confidence", y = "Recall accuracy (strict)", color = "Expertise score", fill = "Expertise score") +
  theme_bw()
```

---

```{r}
j_recall %>% 
  left_join(q_google_demos %>% mutate(subj_num = factor(subj_num)), by = "subj_num") %>% 
  ggplot(aes(x = resp_conf, y = as.integer(acc_recall), color = gender, fill = gender)) + 
  geom_jitter(alpha = 0.3, height = 0.05) + 
  geom_smooth(aes(group = subj_num), method = "glm", method.args = list(family = "binomial"), se = FALSE) +
  labs(x = "Recall confidence", y = "Recall accuracy (strict)", color = "Expertise score", fill = "Expertise score") +
  theme_bw()
```

---

```{r}
j_recall %>% 
  ggplot(aes(x = rt_end_recall, y = as.integer(acc_recall), color = j_score, fill = j_score)) + 
  geom_jitter(alpha = 0.3, height = 0.05) + 
  geom_smooth(aes(group = subj_num), method = "glm", method.args = list(family = "binomial"), alpha = 0.05) + 
  scale_color_viridis_c() +
  scale_fill_viridis_c() +
  labs(x = "Recall submit RT", y = "Recall accuracy (strict)", color = "Expertise score", fill = "Expertise score") +
  theme_bw()
```

---

```{r}
j_recall %>% 
  ggplot(aes(x = rt_end_recall, y = resp_conf, color = j_score, fill = j_score)) + 
  geom_jitter(alpha = 0.3, height = 0.05) + 
  geom_smooth(aes(group = subj_num), method = "lm", alpha = 0.05) + 
  scale_color_viridis_c() +
  scale_fill_viridis_c() +
  labs(x = "Recall submit RT", y = "Recall confidence", color = "Expertise score", fill = "Expertise score") +
  theme_bw()
```

---

### Metamemory state counts

```{r}
j_meta %>%
  filter(!is.na(resp)) %>% 
  left_join(j_recall %>% select(subj_num, answer, acc_recall)) %>% 
  mutate(acc_recall = recode(as.character(acc_recall),
                             `FALSE` = "incorrect",
                             `TRUE` = "correct")) %>% 
  ggplot(aes(x = subj_num, fill = resp)) +
  geom_bar() +
  scale_x_discrete(guide = guide_axis(angle = 30)) +
  facet_grid(~acc_recall) +
  labs(x = "Participant", y = "# facts labeled with that memory state", fill = "Memory state") +
  theme_bw()
```

---

### Metamemory state relative percentages

```{r}
j_meta %>%
  left_join(j_recall %>% select(subj_num, answer, acc_recall)) %>% 
  mutate(acc_recall = recode(as.character(acc_recall),
                             `FALSE` = "incorrect",
                             `TRUE` = "correct")) %>% 
  count(subj_num, acc_recall, resp) %>% 
  group_by(subj_num, acc_recall) %>% 
  mutate(prob = n / sum(n)) %>% 
  ggplot(aes(x = subj_num, y = prob, fill = resp)) +
  geom_col() +
  scale_x_discrete(guide = guide_axis(angle = 30)) +
  facet_grid(~acc_recall) +
  labs(x = "Participant", y = "Percent facts labeled with that memory state", fill = "Memory state") +
  theme_bw()
```

---

### Metamemory descriptions: Guess

```{r, results = "asis"}
j_meta_desc %>%
  filter(question == "Guess") %>% 
  arrange(subj_num) %>% 
  pull(resp) %>% 
  paste0(">", .) %>% 
  cat(sep = "\n\n")
```

---

### Metamemory descriptions: Just Know

```{r, results = "asis"}
j_meta_desc %>%
  filter(question == "Just Know") %>% 
  arrange(subj_num) %>% 
  pull(resp) %>% 
  paste0(">", .) %>% 
  cat(sep = "\n\n")
```

---

### Metamemory descriptions: Related Memory

```{r, results = "asis"}
j_meta_desc %>%
  filter(question == "Related Memory") %>% 
  arrange(subj_num) %>% 
  pull(resp) %>% 
  paste0(">", .) %>% 
  cat(sep = "\n\n")
```

---

### Metamemory descriptions: Learning Memory

```{r, results = "asis"}
j_meta_desc %>%
  filter(question == "Learning Memory") %>% 
  arrange(subj_num) %>% 
  pull(resp) %>% 
  paste0(">", .) %>% 
  cat(sep = "\n\n")
```

---

class: inverse, middle

# Self-report measures

---

```{r}
q_demos %>%
  select(subj_num, starts_with("trivia"), j_score) %>% 
  mutate(j_score = j_score * 100) %>% 
  pivot_longer(cols = -subj_num, names_to = "var", values_to = "value") %>% 
  ggplot(aes(x = value)) + geom_histogram() + facet_wrap(~ var)
```

---

class: inverse, middle

# Museum task encoding

---

```{r}
encoding %>% 
  ggplot(aes(x = resp, fill = category)) +
  geom_histogram(position = "identity", alpha = 0.5) +
  facet_wrap(~subj_num)
```

---

```{r}
encoding %>% 
  ggplot(aes(x = resp, y = rt, color = j_score, fill = j_score)) +
  geom_point() +
  geom_smooth(aes(group = subj_num), method = "lm", se = FALSE) +
  scale_colour_viridis_c() +
  scale_fill_viridis_c()
```

---

```{r}
encoding %>% 
  mutate(already_knew = fct_relevel(already_knew, "Did not know this before", "Knew some of this before")) %>% 
  ggplot(aes(x = subj_num, fill = already_knew)) +
  geom_bar() +
  scale_x_discrete(guide = guide_axis(angle = 30)) +
  labs(x = "Participant", y = "# facts already knew", fill = "Prior knowledge") +
  theme_bw()
```

---

class: inverse, middle

# Museum task retrieval

---

### Fact recall by question order/chunk type

```{r}
retrieval %>% 
  ggplot(aes(x = retrieval_trial_num, y = acc_recall)) +
  geom_rect(aes(x = NULL, y = NULL, xmin = xmin, xmax = xmax, fill = retrieval_block_type),
            data = retrieval %>%
              select(subj_num, retrieval_block, retrieval_trial_num, retrieval_block_type) %>%
              group_by(subj_num, retrieval_block) %>%
              filter(retrieval_trial_num %in% range(retrieval_trial_num)) %>%
              mutate(xtype = if_else(retrieval_trial_num == min(retrieval_trial_num), "xmin", "xmax")) %>%
              ungroup() %>%
              pivot_wider(names_from = xtype, values_from = retrieval_trial_num) %>% 
              mutate(xmin = xmin - 0.5,
                     xmax = xmax + 0.5),
            ymin = 0, ymax = 1, alpha = 0.5) +
  geom_line() +
  geom_point(aes(color = already_knew)) +
  # scale_colour_viridis_b() +
  scale_y_continuous(n.breaks = 3) +
  facet_grid(subj_num ~ .) +
  labs(x = "Fact recall trial (beginning to end)", y = "Fact recall accuracy", fill = "Chunk type") +
  theme_bw()
```

---

### Fact recall by chunk type

```{r}
r_facts %>% 
  mutate(ep_is_same = recode(ep_is_same,
                                         `0` = "ep diff",
                                         `1` = "ep same"),
                     sem_is_same = recode(sem_is_same,
                                          `0` = "sem diff",
                                          `1` = "sem same")) %>% 
  group_by(subj_num, j_score, ep_is_same, sem_is_same) %>% 
  summarize(mean_acc = mean(acc_recall)) %>% 
  ggplot(aes(x = interaction(ep_is_same, sem_is_same), y = mean_acc,
             group = subj_num,
             color = j_score)) +
  geom_line() +
  geom_point() +
  scale_x_discrete(guide = guide_axis(angle = 30)) +
  scale_color_viridis_c() +
  labs(x = "Chunk type", y = "Percent facts correctly recalled", color = "Expertise score") +
  theme_bw()
```

---

### Fact recall by already knew status

```{r}
retrieval %>% 
  mutate(acc_recall = if_else(acc_recall > 0, 1, 0)) %>% 
  filter(!is.na(already_knew)) %>% 
  ggplot(aes(x = subj_num, fill = factor(acc_recall))) +
  geom_bar() +
  facet_grid(~already_knew)
```

---

```{r}
retrieval %>% 
  mutate(acc_recall = if_else(acc_recall > 0, 1, 0)) %>% 
  filter(!is.na(already_knew)) %>% 
  count(subj_num, acc_recall, already_knew) %>% 
  group_by(subj_num, already_knew) %>% 
  mutate(prob = n / sum(n)) %>% 
  ggplot(aes(x = subj_num, y = prob, fill = factor(acc_recall))) +
  geom_col() +
  facet_grid(~already_knew)
```

---

### Fact recall by ENCODING question order

```{r}
retrieval %>% 
  nest(data = -subj_num) %>% 
  sample_n(10) %>% 
  unnest(data) %>% 
  mutate(encoding_trial_num = if_else(group == "academic", group_trial_num * 2 - 1, group_trial_num * 2)) %>% 
  ggplot(aes(x = retrieval_trial_num, y = acc_recall)) +
  geom_line() +
  geom_point(aes(color = already_knew)) +
  # scale_colour_viridis_b() +
  scale_y_continuous(n.breaks = 3) +
  facet_grid(subj_num ~ .) +
  labs(x = "Encoding trial (beginning to end)", y = "Subsequent fact recall accuracy") +
  theme_bw()
```

---

### Photo recognition

Current setup: A-A' 2AFC with similar lure

```{r}
r_pics %>%
  ggplot(aes(x = factor(subj_num), y = resp, color = j_score)) +
  geom_hline(yintercept = 0, linetype = "dotted") +
  geom_boxplot() +
  scale_colour_viridis_c() +
  guides(color = FALSE) +
  labs(x = "Participant", y = "Accuracy with confidence (positive = correct)") +
  theme_bw()
```

---

### Source memory for which museum

```{r}
r_source %>%
  ggplot(aes(x = factor(subj_num), y = resp, color = j_score)) +
  geom_hline(yintercept = 0, linetype = "dotted") +
  geom_boxplot() +
  scale_colour_viridis_c() +
  guides(color = FALSE) +
  labs(x = "Participant", y = "Accuracy with confidence (positive = correct)") +
  theme_bw()
```

---

class: inverse, middle

# Retrieval between-task relationships

---

### Does _photo_ retrieval predict _fact_ retrieval?

```{r}
retrieval %>% 
  # filter(!is.na(already_knew)) %>% 
  mutate(acc_recall = if_else(acc_recall == 0.5, 1, acc_recall)) %>%
  ggplot(aes(x = resp_pic, y = acc_recall, color = j_score, fill = j_score)) + 
  geom_jitter(alpha = 0.3, height = 0.05) + 
  geom_smooth(aes(group = subj_num), method = "glm", method.args = list(family = "binomial"), alpha = 0.05) + 
  scale_color_viridis_c() +
  scale_fill_viridis_c() +
  # facet_grid(~already_knew) +
  labs(x = "Photo recognition accuracy", y = "Fact recall accuracy (generous)") +
  theme_bw()
```

---

### Does _source_ retrieval predict _fact_ retrieval?

```{r}
retrieval %>% 
  mutate(acc_recall = if_else(acc_recall == 0.5, 1, acc_recall)) %>%
  ggplot(aes(x = resp_source, y = acc_recall, color = j_score, fill = j_score)) + 
  geom_jitter(alpha = 0.3, height = 0.05) + 
  geom_smooth(aes(group = subj_num), method = "glm", method.args = list(family = "binomial"), alpha = 0.05) + 
  scale_color_viridis_c() +
  scale_fill_viridis_c() +
  labs(x = "Source recognition accuracy", y = "Fact recall accuracy (generous)") +
  theme_bw()
```

---

### Does _source_ retrieval predict _photo_ retrieval?

```{r}
retrieval %>% 
  ggplot(aes(x = resp_source, y = resp_pic, color = j_score, fill = j_score)) + 
  geom_point(alpha = 0.3) + 
  geom_smooth(aes(group = subj_num), method = "lm", se = FALSE) + 
  scale_color_viridis_c() +
  scale_fill_viridis_c() +
  labs(x = "Source recognition accuracy", y = "Photo recognition accuracy") +
  theme_bw()
```

---

### Does interest at encoding predict _fact_ retrieval?

```{r}
retrieval %>% 
  mutate(acc_recall = if_else(acc_recall == 0.5, 1, acc_recall)) %>%
  ggplot(aes(x = interest, y = acc_recall, color = j_score, fill = j_score)) + 
  geom_jitter(alpha = 0.3, height = 0.05) + 
  geom_smooth(aes(group = subj_num), method = "glm", method.args = list(family = "binomial"), alpha = 0.05) + 
  scale_color_viridis_c() +
  scale_fill_viridis_c() +
  labs(x = "Interest in fact at encoding", y = "Fact recall accuracy (generous)") +
  theme_bw()
```

---

### Does interest at encoding predict _photo_ retrieval?

```{r}
retrieval %>% 
  ggplot(aes(x = interest, y = resp_pic, color = j_score, fill = j_score)) + 
  geom_point(alpha = 0.3) + 
  geom_smooth(aes(group = subj_num), method = "lm", se = FALSE) + 
  scale_color_viridis_c() +
  scale_fill_viridis_c() +
  labs(x = "Interest in fact at encoding", y = "Photo recognition accuracy") +
  theme_bw()
```

---

### Does interest at encoding predict _source_ retrieval?

```{r}
retrieval %>% 
  ggplot(aes(x = interest, y = resp_source, color = j_score, fill = j_score)) + 
  geom_point(alpha = 0.3) + 
  geom_smooth(aes(group = subj_num), method = "lm", se = FALSE) + 
  scale_color_viridis_c() +
  scale_fill_viridis_c() +
  labs(x = "Interest in fact at encoding", y = "Source recognition accuracy") +
  theme_bw()
```

---

class: inverse, middle

# STOP! Dirty graphs ahead

---

```{r}
retrieval %>% 
  ggplot(aes(x = log(rt_pic), y = resp_pic)) + 
  geom_point() + 
  geom_smooth(method = "lm") + 
  facet_wrap(~subj_num, scales = "free_x")
```

```{r}
retrieval %>% 
  ggplot(aes(x = log(rt_source), y = resp_source)) + 
  geom_point() + 
  geom_smooth(method = "lm") + 
  facet_wrap(~subj_num, scales = "free_x")
```