### 2021-10-12 slides/data_checking.Rmd
needs to be cleaned up but started

---
title: "Exploratory data analysis"
subtitle: "J-Study"
author: "Monica Thieu & Lauren Wilkins"
institute: "Columbia University Dept of Psychology"
date: "`r Sys.Date()`"
output:
  xaringan::moon_reader:
    css: ["theme.css", "fonts.css"]
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, dev = "svglite", message = FALSE, warning = FALSE, fig.asp = 3/4)
require(tidyverse)
require(magrittr)

demos_ids <- read_csv(here::here("ignore", "data", "demos_ids.csv"))
q_google_demos <- read_csv(here::here("ignore", "data", "q_google_demos.csv")) %>% 
  right_join(demos_ids, by = "demographics_id") %>% 
  mutate(subj_num = factor(subj_num))

j_recall <- read_csv(here::here("ignore", "data", "task_jeopardy_recall.csv")) %>% 
  mutate(resp_conf = if_else(is.na(resp_recall), NA_real_, resp_conf))

j_exp <- j_recall %>% 
  group_by(subj_num) %>% 
  summarize(j_score = mean(acc_recall))

j_recall %<>%
  left_join(j_exp, by = "subj_num") %>% 
  mutate(subj_num = fct_reorder(as.character(subj_num), j_score))

j_meta <- read_csv(here::here("ignore", "data", "task_jeopardy_meta_states.csv")) %>% 
  left_join(j_exp, by = "subj_num") %>% 
  mutate(subj_num = fct_reorder(as.character(subj_num), j_score),
         resp = fct_relevel(resp, "Learning Memory", after = Inf))
j_meta_desc <- read_csv(here::here("ignore", "data", "task_jeopardy_meta_descriptions.csv")) %>% 
  left_join(j_exp, by = "subj_num") %>% 
  mutate(subj_num = fct_reorder(as.character(subj_num), j_score))
encoding <- read_csv(here::here("ignore", "data", "task_encoding.csv")) %>% 
  left_join(j_exp, by = "subj_num") %>% 
  mutate(subj_num = fct_reorder(as.character(subj_num), j_score),
         already_knew = fct_recode(already_knew,
                               none = "Did not know this before",
                               some = "Knew some of this before",
                               all = "Knew all of this before"))
r_facts <- read_csv(here::here("ignore", "data",  "task_retrieval_facts.csv")) %>% 
  left_join(j_exp, by = "subj_num") %>% 
  mutate(subj_num = fct_reorder(as.character(subj_num), j_score))
r_pics <- read_csv(here::here("ignore", "data", "task_retrieval_pics.csv")) %>% 
  left_join(j_exp, by = "subj_num") %>% 
  mutate(subj_num = fct_reorder(as.character(subj_num), j_score))
r_source <- read_csv(here::here("ignore", "data", "task_retrieval_source.csv")) %>% 
  left_join(j_exp, by = "subj_num") %>% 
  mutate(subj_num = fct_reorder(as.character(subj_num), j_score))
q_trivia_demos  <- read_csv(here::here("ignore", "data", "q_trivia_demos.csv")) %>% 
  left_join(j_exp, by = "subj_num") %>% 
  mutate(subj_num = fct_reorder(as.character(subj_num), j_score))

q_posttask  <- read_csv(here::here("ignore", "data", "q_posttask.csv"))

retrieval <- r_facts %>%
  select(subj_num, j_score, group, group_trial_num, retrieval_block, ep_is_same, sem_is_same, retrieval_trial_num, category, acc_recall, rt_start_recall = rt_start, rt_end_recall = rt_end) %>% 
  left_join(r_pics %>% 
              select(subj_num, group, group_trial_num, category, resp_pic = resp, rt_pic = rt)) %>% 
  left_join(r_source %>%
              select(subj_num, group, group_trial_num, category, resp_source = resp, rt_source = rt)) %>%
  left_join(encoding %>% 
              select(subj_num, group, group_trial_num, category, interest = resp, already_knew)) %>% 
  mutate(retrieval_block_type = case_when(retrieval_block <= 4 ~ "ep_same_sem_same",
                                          retrieval_block <= 8 ~ "ep_same_sem_diff",
                                          retrieval_block <= 12 ~ "ep_diff_sem_same",
                                          retrieval_block <= 16 ~ "ep_diff_sem_diff",
                                          TRUE ~ NA_character_),
         encoding_trial_num = if_else(group == "academic",
                                      group_trial_num * 2 - 1,
                                      group_trial_num * 2))
```

class: inverse, middle

# Demographics

---

### What is the age and gender breakdown?

```{r}
q_google_demos %>% 
  ggplot(aes(x = age, fill = gender)) +
  geom_histogram(position = "identity", alpha = 0.5, binwidth = 2)
```

---

class: inverse, middle

# Trivia expertise assessment

---

### What is our sample spread of trivia expertise?

```{r}
j_exp %>% 
  ggplot(aes(x = j_score)) +
  geom_histogram() +
  geom_vline(xintercept = 0.7, linetype = "dotted")
```

---

### Do Ps accurately assess their own trivia skill?

```{r}
q_trivia_demos %>% 
  left_join(q_google_demos, by = "subj_num") %>% 
  ggplot(aes(x = trivia_skill_self, y = j_score)) +
  # geom_abline(slope = .01, intercept = 0, linetype = "dotted") +
  geom_vline(xintercept = 75, linetype = "dotted") +
  geom_hline(yintercept = 0.7, linetype = "dotted") +
  geom_point(aes(color = gender), size = 2, alpha = 0.7) +
  geom_smooth(method = "lm", formula = y ~ x) +
  expand_limits(y = 0) +
  labs(x = "Self-reported trivia expertise (relative to 'average person')",
       y = "Expertise score on Jeopardy audition test") +
  theme_bw()
```

---

### Trivia recall by confidence

```{r}
j_recall %>% 
  ggplot(aes(x = resp_conf, y = as.integer(acc_recall), color = j_score, fill = j_score)) + 
  geom_jitter(alpha = 0.3, height = 0.05) + 
  geom_smooth(aes(group = subj_num), method = "glm", method.args = list(family = "binomial"), se = FALSE) + 
  scale_color_viridis_c() +
  scale_fill_viridis_c() +
  labs(x = "Recall confidence",
       y = "Recall accuracy (strict)",
       color = "Expertise score",
       fill = "Expertise score") +
  theme_bw()
```

---

### Better experts tend to be more confident

This effect does not necessarily imply that better experts are _overconfident._ Rather, it is consistent with better experts tending to have higher latent knowledge for each question, and thus being likely to be more confident overall, even when the answer turns out to be wrong.

```{r}
j_recall %>% 
  group_by(subj_num, j_score) %>% 
  summarize(conf_q500 = median(resp_conf, na.rm = TRUE),
            conf_q250 = quantile(resp_conf, .25, na.rm = TRUE),
            conf_q750 = quantile(resp_conf, .75, na.rm = TRUE),
            conf_q025 = quantile(resp_conf, .025, na.rm = TRUE),
            conf_q975 = quantile(resp_conf, .975, na.rm = TRUE)) %>% 
  ggplot(aes(x = subj_num, y = conf_q500, color = j_score)) +
  # geom_errorbar(aes(ymin = conf_q025, ymax = conf_q975), width = 0, size = 0.5) +
  geom_errorbar(aes(ymin = conf_q250, ymax = conf_q750), width = 0, size = 1) +
  geom_point(size = 3) +
  coord_flip() +
  scale_color_viridis_c() +
  theme_bw() +
  theme(axis.text.y = element_blank())
```

---

---

```{r}
j_recall %>% 
  ggplot(aes(x = rt_end_recall, y = as.integer(acc_recall), color = j_score, fill = j_score)) + 
  geom_jitter(alpha = 0.3, height = 0.05) + 
  geom_smooth(aes(group = subj_num), method = "glm", method.args = list(family = "binomial"), se = FALSE) + 
  scale_color_viridis_c() +
  scale_fill_viridis_c() +
  labs(x = "Recall submit RT", y = "Recall accuracy (strict)", color = "Expertise score", fill = "Expertise score") +
  theme_bw()
```

---

### Better experts tend to respond faster

Again, this is consistent with an across-participants tendency to respond faster for subsequently correct answers. Better experts who are giving more correct answers are thus more likely to have shorter RTs.

```{r}
j_recall %>% 
  group_by(subj_num, j_score) %>% 
  summarize(rt_q500 = median(rt_end_recall, na.rm = TRUE),
            rt_q250 = quantile(rt_end_recall, .25, na.rm = TRUE),
            rt_q750 = quantile(rt_end_recall, .75, na.rm = TRUE),
            rt_q025 = quantile(rt_end_recall, .025, na.rm = TRUE),
            rt_q975 = quantile(rt_end_recall, .975, na.rm = TRUE),
            rt_mean = mean(rt_end_recall, na.rm = TRUE),
            rt_sd = sd(rt_end_recall, na.rm = TRUE)) %>% 
  ggplot(aes(x = subj_num, y = rt_mean, color = j_score)) +
  geom_errorbar(aes(ymin = rt_mean - rt_sd/2, ymax = rt_mean + rt_sd/2), width = 0, size = 1) +
  geom_point(size = 3) +
  coord_flip() +
  scale_color_viridis_c() +
  theme_bw() +
  theme(axis.text.y = element_blank())
```

---

```{r}
j_recall %>% 
  ggplot(aes(x = rt_end_recall, y = resp_conf, color = j_score, fill = j_score)) + 
  geom_jitter(alpha = 0.3, height = 0.05) + 
  geom_smooth(aes(group = subj_num), method = "lm", alpha = 0.05) + 
  scale_color_viridis_c() +
  scale_fill_viridis_c() +
  labs(x = "Recall submit RT", y = "Recall confidence", color = "Expertise score", fill = "Expertise score") +
  theme_bw()
```

---

### Metamemory state counts

```{r}
j_meta %>%
  filter(!is.na(resp)) %>% 
  left_join(j_recall %>% select(subj_num, answer, acc_recall)) %>% 
  mutate(acc_recall = recode(as.character(acc_recall),
                             `FALSE` = "incorrect",
                             `TRUE` = "correct")) %>% 
  ggplot(aes(x = subj_num, fill = resp)) +
  geom_bar() +
  scale_x_discrete(guide = guide_axis(angle = 30)) +
  facet_grid(~acc_recall) +
  labs(x = "Participant", y = "# facts labeled with that memory state", fill = "Memory state") +
  theme_bw()
```

---

### Metamemory state relative percentages

```{r}
j_meta %>%
  left_join(j_recall %>% select(subj_num, answer, acc_recall)) %>% 
  mutate(acc_recall = recode(as.character(acc_recall),
                             `FALSE` = "incorrect",
                             `TRUE` = "correct")) %>% 
  count(subj_num, acc_recall, resp) %>% 
  group_by(subj_num, acc_recall) %>% 
  mutate(prob = n / sum(n)) %>% 
  ggplot(aes(x = subj_num, y = prob, fill = resp)) +
  geom_col() +
  scale_x_discrete(guide = guide_axis(angle = 30)) +
  facet_grid(~acc_recall) +
  labs(x = "Participant", y = "Percent facts labeled with that memory state", fill = "Memory state") +
  theme_bw()
```

---

### Metamemory descriptions: Guess

```{r, results = "asis"}
j_meta_desc %>%
  filter(question == "Guess") %>% 
  arrange(subj_num) %>% 
  pull(resp) %>% 
  paste0(">", .) %>% 
  cat(sep = "\n\n")
```

---

### Metamemory descriptions: Just Know

```{r, results = "asis"}
j_meta_desc %>%
  filter(question == "Just Know") %>% 
  arrange(subj_num) %>% 
  pull(resp) %>% 
  paste0(">", .) %>% 
  cat(sep = "\n\n")
```

---

### Metamemory descriptions: Related Memory

```{r, results = "asis"}
j_meta_desc %>%
  filter(question == "Related Memory") %>% 
  arrange(subj_num) %>% 
  pull(resp) %>% 
  paste0(">", .) %>% 
  cat(sep = "\n\n")
```

---

### Metamemory descriptions: Learning Memory

```{r, results = "asis"}
j_meta_desc %>%
  filter(question == "Learning Memory") %>% 
  arrange(subj_num) %>% 
  pull(resp) %>% 
  paste0(">", .) %>% 
  cat(sep = "\n\n")
```

---

class: inverse, middle

# Self-report measures

---

```{r}
q_trivia_demos %>%
  select(subj_num, starts_with("trivia"), j_score) %>% 
  mutate(j_score = j_score * 100) %>% 
  pivot_longer(cols = -subj_num, names_to = "var", values_to = "value") %>% 
  ggplot(aes(x = value)) + geom_histogram() + facet_wrap(~ var)
```

---

class: inverse, middle

# Museum task encoding

---

```{r}
encoding %>% 
  ggplot(aes(x = resp, fill = category)) +
  geom_histogram(position = "identity", binwidth = 5, alpha = 0.5)
```

---

```{r}
encoding %>% 
  ggplot(aes(x = j_score, y = resp, color = category, fill = category)) +
  geom_jitter(alpha = 0.5) +
  geom_smooth(formula = y ~ x, method = "lm")
```

---

```{r}
encoding %>% 
  ggplot(aes(x = resp, y = rt, color = j_score, fill = j_score)) +
  geom_point() +
  geom_smooth(aes(group = subj_num), method = "lm", se = FALSE) +
  scale_colour_viridis_c() +
  scale_fill_viridis_c()
```

---

```{r}
encoding %>% 
  mutate(already_knew = fct_relevel(already_knew, "none", "some")) %>% 
  ggplot(aes(x = subj_num, fill = already_knew)) +
  geom_bar() +
  scale_x_discrete(guide = guide_axis(angle = 30)) +
  labs(x = "Participant", y = "# facts already knew", fill = "Prior knowledge") +
  theme_bw()
```

---

class: inverse, middle

# Museum task retrieval

---

### Fact recall by question order/chunk type

```{r}
retrieval %>% 
  ggplot(aes(x = retrieval_trial_num, y = acc_recall)) +
  geom_rect(aes(x = NULL, y = NULL, xmin = xmin, xmax = xmax, fill = retrieval_block_type),
            data = retrieval %>%
              select(subj_num, retrieval_block, retrieval_trial_num, retrieval_block_type) %>%
              group_by(subj_num, retrieval_block) %>%
              filter(retrieval_trial_num %in% range(retrieval_trial_num)) %>%
              mutate(xtype = if_else(retrieval_trial_num == min(retrieval_trial_num), "xmin", "xmax")) %>%
              ungroup() %>%
              pivot_wider(names_from = xtype, values_from = retrieval_trial_num) %>% 
              mutate(xmin = xmin - 0.5,
                     xmax = xmax + 0.5),
            ymin = 0, ymax = 1, alpha = 0.5) +
  geom_line() +
  geom_point(aes(color = already_knew)) +
  # scale_colour_viridis_b() +
  scale_y_continuous(n.breaks = 3) +
  facet_grid(subj_num ~ .) +
  labs(x = "Fact recall trial (beginning to end)", y = "Fact recall accuracy", fill = "Chunk type") +
  theme_bw()
```

---

### Fact recall by chunk type

```{r}
r_facts %>% 
  mutate(ep_is_same = recode(ep_is_same,
                                         `0` = "ep diff",
                                         `1` = "ep same"),
                     sem_is_same = recode(sem_is_same,
                                          `0` = "sem diff",
                                          `1` = "sem same")) %>% 
  group_by(subj_num, j_score, ep_is_same, sem_is_same) %>% 
  summarize(mean_acc = mean(acc_recall)) %>% 
  ggplot(aes(x = interaction(ep_is_same, sem_is_same), y = mean_acc,
             group = subj_num,
             color = j_score)) +
  geom_line() +
  geom_point() +
  scale_x_discrete(guide = guide_axis(angle = 30)) +
  scale_color_viridis_c() +
  labs(x = "Chunk type", y = "Percent facts correctly recalled", color = "Expertise score") +
  theme_bw()
```

---

### Fact recall by already knew status

```{r}
retrieval %>% 
  mutate(acc_recall = if_else(acc_recall > 0, 1, 0)) %>% 
  filter(!is.na(already_knew)) %>% 
  ggplot(aes(x = subj_num, fill = factor(acc_recall))) +
  geom_bar() +
  facet_grid(~already_knew)
```

---

```{r}
retrieval %>% 
  mutate(acc_recall = if_else(acc_recall > 0, 1, 0)) %>% 
  filter(!is.na(already_knew)) %>% 
  count(subj_num, acc_recall, already_knew) %>% 
  group_by(subj_num, already_knew) %>% 
  mutate(prob = n / sum(n)) %>% 
  ggplot(aes(x = subj_num, y = prob, fill = factor(acc_recall))) +
  geom_col() +
  facet_grid(~already_knew)
```

---

### Fact recall by ENCODING question order

```{r}
retrieval %>% 
  nest(data = -subj_num) %>% 
  sample_n(10) %>% 
  unnest(data) %>% 
  mutate(encoding_trial_num = if_else(group == "academic", group_trial_num * 2 - 1, group_trial_num * 2)) %>% 
  ggplot(aes(x = retrieval_trial_num, y = acc_recall)) +
  geom_line() +
  geom_point(aes(color = already_knew)) +
  # scale_colour_viridis_b() +
  scale_y_continuous(n.breaks = 3) +
  facet_grid(subj_num ~ .) +
  labs(x = "Encoding trial (beginning to end)", y = "Subsequent fact recall accuracy") +
  theme_bw()
```

---

### Photo recognition

Current setup: A-A' 2AFC with similar lure

```{r}
r_pics %>%
  ggplot(aes(x = factor(subj_num), y = resp, color = j_score)) +
  geom_hline(yintercept = 0, linetype = "dotted") +
  geom_boxplot() +
  scale_colour_viridis_c() +
  guides(color = FALSE) +
  labs(x = "Participant", y = "Accuracy with confidence (positive = correct)") +
  theme_bw()
```

---

### Source memory for which museum

```{r}
r_source %>%
  ggplot(aes(x = factor(subj_num), y = resp, color = j_score)) +
  geom_hline(yintercept = 0, linetype = "dotted") +
  geom_boxplot() +
  scale_colour_viridis_c() +
  guides(color = FALSE) +
  labs(x = "Participant", y = "Accuracy with confidence (positive = correct)") +
  theme_bw()
```

---

class: inverse, middle

# Retrieval between-task relationships

---

### Does _photo_ retrieval predict _fact_ retrieval?

```{r}
retrieval %>% 
  filter(already_knew == "none") %>% 
  mutate(acc_recall = if_else(acc_recall == 0.5, 1, acc_recall)) %>%
  ggplot(aes(x = resp_pic, y = acc_recall, color = j_score, fill = j_score)) + 
  geom_jitter(alpha = 0.3, height = 0.05) + 
  geom_smooth(aes(group = subj_num), method = "glm", method.args = list(family = "binomial"), alpha = 0.02) + 
  scale_color_viridis_c() +
  scale_fill_viridis_c() +
  # facet_grid(~already_knew) +
  labs(x = "Photo recognition accuracy", y = "Fact recall accuracy (generous)") +
  theme_bw()
```

---

```{r}
retrieval %>% 
  filter(already_knew == "none") %>% 
  mutate(acc_recall = if_else(acc_recall == 0.5, 1, acc_recall)) %>%
  ggplot(aes(x = resp_pic, y = acc_recall, color = cut_number(j_score, 3), fill = cut_number(j_score, 3))) + 
  geom_jitter(alpha = 0.3, height = 0.05) + 
  geom_smooth( method = "glm", method.args = list(family = "binomial"), alpha = 0.2) + 
  labs(x = "Photo recognition accuracy", y = "Fact recall accuracy (generous)") +
  theme_bw()
```

---

```{r}
retrieval %>% 
  filter(already_knew == "none") %>% 
  mutate(acc_recall = as.numeric(acc_recall > 0),
         resp_pic = resp_pic / 100,
         j_score = j_score - 0.5) %>% 
  lme4::glmer(acc_recall ~ resp_pic * j_score + (1 | subj_num), data = ., family = "binomial") %>% 
  summary()
```

---

### Does _source_ retrieval predict _fact_ retrieval?

```{r}
retrieval %>% 
  mutate(acc_recall = if_else(acc_recall == 0.5, 1, acc_recall)) %>%
  ggplot(aes(x = resp_source, y = acc_recall, color = j_score, fill = j_score)) + 
  geom_jitter(alpha = 0.3, height = 0.05) + 
  geom_smooth(aes(group = subj_num), method = "glm", method.args = list(family = "binomial"), alpha = 0.05) + 
  scale_color_viridis_c() +
  scale_fill_viridis_c() +
  labs(x = "Source recognition accuracy", y = "Fact recall accuracy (generous)") +
  theme_bw()
```

---

```{r}
retrieval %>% 
  filter(already_knew == "none") %>% 
  mutate(acc_recall = if_else(acc_recall == 0.5, 1, acc_recall),
         j_score_binary = j_score > 0.72) %>%
  ggplot(aes(x = resp_source, y = acc_recall, color = cut_number(j_score, 4), fill = cut_number(j_score, 4))) + 
  geom_jitter(alpha = 0.3, height = 0.05) + 
  geom_smooth( method = "glm", method.args = list(family = "binomial"), alpha = 0.2) + 
  labs(x = "Source recognition accuracy", y = "Fact recall accuracy (generous)") +
  theme_bw()
```

---

```{r}
retrieval %>% 
  filter(already_knew == "none") %>% 
  mutate(acc_recall = as.numeric(acc_recall > 0),
         resp_source = resp_source / 100,
         j_score = j_score - 0.5) %>% 
  lme4::glmer(acc_recall ~ resp_source * j_score + (1 | subj_num), data = ., family = "binomial") %>% 
  summary()
```

---

### Does _source_ retrieval predict _photo_ retrieval?

```{r}
retrieval %>% 
  ggplot(aes(x = resp_source, y = resp_pic, color = j_score, fill = j_score)) + 
  geom_point(alpha = 0.3) + 
  geom_smooth(aes(group = subj_num), method = "lm", se = FALSE) + 
  scale_color_viridis_c() +
  scale_fill_viridis_c() +
  labs(x = "Source recognition accuracy", y = "Photo recognition accuracy") +
  theme_bw()
```

---

```{r}
retrieval %>% 
  filter(already_knew == "none") %>% 
  ggplot(aes(x = resp_source, y = resp_pic, color = cut_number(j_score, 5), fill = cut_number(j_score, 5))) + 
  geom_point(alpha = 0.1) + 
  geom_smooth( method = "lm", alpha = 0.2) + 
  labs(x = "Source recognition accuracy", y = "Photo recognition accuracy") +
  theme_bw()
```

---

### Does interest at encoding predict _fact_ retrieval?

```{r}
retrieval %>% 
  mutate(acc_recall = if_else(acc_recall == 0.5, 1, acc_recall)) %>%
  ggplot(aes(x = interest, y = acc_recall, color = j_score, fill = j_score)) + 
  geom_jitter(alpha = 0.3, height = 0.05) + 
  geom_smooth(aes(group = subj_num), method = "glm", method.args = list(family = "binomial"), alpha = 0.05) + 
  scale_color_viridis_c() +
  scale_fill_viridis_c() +
  labs(x = "Interest in fact at encoding", y = "Fact recall accuracy (generous)") +
  theme_bw()
```

---

```{r}
retrieval %>% 
  filter(already_knew == "none") %>% 
  mutate(acc_recall = if_else(acc_recall == 0.5, 1, acc_recall)) %>%
  ggplot(aes(x = interest, y = acc_recall, color = cut_number(j_score, 4), fill = cut_number(j_score, 4))) + 
  geom_jitter(alpha = 0.3, height = 0.05) + 
  geom_smooth( method = "glm", method.args = list(family = "binomial"), alpha = 0.2) + 
  labs(x = "Interest in fact at encoding", y = "Fact recall accuracy (generous)") +
  theme_bw()
```

---

### Does interest at encoding predict _photo_ retrieval?

```{r}
retrieval %>% 
  # filter(already_knew == "none") %>% 
  ggplot(aes(x = interest, y = resp_pic, color = cut_number(j_score, 4), fill = cut_number(j_score, 4))) + 
  geom_point(alpha = 0.1) + 
  geom_smooth(method = "lm", alpha = 0.2) + 
  labs(x = "Interest in fact at encoding", y = "Photo recognition accuracy") +
  theme_bw()
```

---

### Does interest at encoding predict _source_ retrieval?

```{r}
retrieval %>% 
  filter(already_knew == "none") %>% 
  ggplot(aes(x = interest, y = resp_source, color = cut_number(j_score, 4), fill = cut_number(j_score, 4))) + 
  geom_point(alpha = 0.1) + 
  geom_smooth(method = "lm", alpha = 0.2) + 
  labs(x = "Interest in fact at encoding", y = "Source recognition accuracy") +
  theme_bw()
```

---

class: inverse, middle

# STOP! Dirty graphs ahead

---

```{r}
retrieval %>% 
  ggplot(aes(x = log(rt_pic), y = resp_pic)) + 
  geom_point() + 
  geom_smooth(method = "lm") + 
  facet_wrap(~subj_num, scales = "free_x")
```

```{r}
retrieval %>% 
  ggplot(aes(x = log(rt_source), y = resp_source)) + 
  geom_point() + 
  geom_smooth(method = "lm") + 
  facet_wrap(~subj_num, scales = "free_x")
```