### 2022-10-13 slides/results.Rmd
catching up on much older work history

---
title: "Preliminary (quick and dirty) results"
subtitle: "J-Study"  
author: 
  - "Monica Thieu"
  - "Lauren Wilkins"
institute: "Columbia University Dept of Psychology"
date: 'Updated: `r Sys.Date()`'
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    nature:
      slideNumberFormat: "%current%"
      highlightStyle: github
      highlightLines: true
      ratio: 16:9
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(
  dev = "svglite", fig.width=9, fig.height=3.5, fig.retina=3,
  out.width = "100%",
  cache = FALSE,
  echo = FALSE,
  message = FALSE, 
  warning = FALSE,
  hiline = TRUE
)

require(xaringanthemer)
source(here::here("R", "get_cleaned_data.R"))
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
style_mono_accent(
  base_color = "#0046a6",
  text_font_google = google_font("Lora"),
  header_font_google = google_font("Nunito"),
  code_font_google = google_font("Fira Code")
)
```

class: inverse, middle

# Proposed main figures

---

### _Photo_ recognition predicts _fact_ recall, slightly more so for experts

```{r}
retrieval %>% 
  filter(already_knew == "none") %>% 
  mutate(across(c(resp_pic, resp_source), ~if_else(. > 0, 1L, 0L)),
         acc_recall = if_else(acc_recall == 0.5, 1, acc_recall)) %>%
  group_by(subj_num, j_score, resp_pic) %>%
  summarize(acc_recall = mean(acc_recall)) %>% 
  ggplot(aes(x = resp_pic, y = acc_recall, color = fct_rev(cut_number(j_score, 3)), fill = fct_rev(cut_number(j_score, 3)))) + 
  geom_line(aes(group = subj_num), alpha = 0.1) +
  geom_smooth(method = "lm", alpha = 0.2) + 
  scale_x_continuous(breaks = 0:1) +
  scale_color_viridis_d(begin = 0.2, end = 0.9, direction = -1) +
  scale_fill_viridis_d(begin = 0.2, end = 0.9, direction = -1) +
  labs(x = "Photo recognition accuracy",
       y = "Fact recall accuracy (generous)",
       color = "Expertise tertile",
       fill = "Expertise tertile") +
  theme_bw()
```

---

class: inverse, middle

# Demographics

---

### More men than women, but age does not appear correlated with gender

```{r}
q_google_demos %>% 
  left_join(q_google_demos %>%
              count(gender) %>%
              mutate(gender_n = glue::glue("{gender} (n = {n})")) %>%
              select(-n),
            by = "gender") %>% 
  ggplot(aes(x = age, fill = gender_n)) +
  geom_histogram(position = "identity", alpha = 0.5, binwidth = 2) +
  scale_y_continuous(breaks = scales::breaks_extended(6))
```

---

### What is our sample spread of trivia expertise?

```{r}
j_exp %>% 
  ggplot(aes(x = j_score)) +
  geom_histogram(binwidth = .05) +
  geom_vline(xintercept = 0.7, linetype = "dotted") +
  annotate("text",
           x = .67,
           y = 20,
           label = "Approx cutoff for Jeopardy callback",
           hjust = 1) +
  labs(x = "Expertise score (fraction correct out of 50 questions)",
       y = "count")
```

---

class: inverse, middle

# Encoding

---

### Trivia experts tend to know more encoding facts already

```{r}
encoding %>% 
  group_by(subj_num, j_score, group) %>% 
  summarize(pct_novel = sum(!is.na(already_knew) & already_knew == "none")) %>% 
  ggplot(aes(x = j_score, y = pct_novel, color = group, fill = group)) +
  geom_hline(yintercept = 40, linetype = "dotted") +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(x = "Trivia expertise", y = "# museum facts rated as novel") +
  theme_bw()
```

---

### Trivia experts might be more interested at encoding, but not by much

```{r}
this_cor <- encoding %>%
  group_by(subj_num, j_score) %>% 
  summarize(interest = median(resp)) %$% 
  cor(j_score, interest) %>% 
  round(2)
```

```{r}
encoding %>% 
  group_by(subj_num, j_score) %>% 
  summarize(resp_median = median(resp),
            q25 = quantile(resp, .25),
            q75 = quantile(resp, .75)) %>% 
  ggplot(aes(x = j_score, y = resp_median)) +
  geom_errorbar(aes(ymin = q25, ymax = q75), width = 0, alpha = 0.5) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(x = "Expertise score",
       y = "P's median interest (across facts)") +
  theme_bw()
```

---

### Trivia experts might be more interested at encoding, but not by much

Quick and dirty maximum-likelihood linear regression

Formula: `interest ~ j_score + already_knew + (1 | subj_num)`

```{r}
encoding %>% 
  mutate(j_score = j_score - 0.7) %>% 
  lme4::lmer(resp ~ j_score + already_knew + (1 | subj_num), data = .) %>% 
  broom.mixed::tidy() %>%
  filter(effect == "fixed") %>% 
  select(term:statistic) %>% 
  mutate(across(where(is.numeric), signif, digits = 3)) %>% 
  knitr::kable()
```

---

class: inverse, middle

# Retrieval

---

### Recall is higher when people already know the facts (duh)

Accordingly, all subsequent results involving fact retrieval are shown only for facts rated as novel at encoding.

Excluding facts known at encoding should not introduce collider bias into the retrieval results, because recall cannot cause prior knowledge at encoding.

```{r}
retrieval %>% 
  filter(!is.na(already_knew)) %>% 
  group_by(subj_num, already_knew) %>% 
  summarize(acc_recall = mean(acc_recall)) %>% 
  ggplot(aes(x = already_knew, y = acc_recall)) +
  geom_violin() +
  labs(x = "Fact prior knowledge at encoding", y = "% correct recall") +
  theme_bw()
```

---

### Trivia experts show better recall for novel facts

```{r}
retrieval %>% 
  filter(already_knew == "none") %>% 
  group_by(subj_num, j_score) %>% 
  summarize(acc_recall = mean(acc_recall)) %>% 
  ggplot(aes(x = j_score, y = acc_recall)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(x = "Expertise score", y = "% correct recall for novel facts") +
  theme_bw()
```

---

### Trivia experts' fact recall is _not_ driven by interest at encoding

.pull-left[

Formula: `acc_recall ~ j_score + (1 | subj_num)`

```{r}
retrieval %>% 
  filter(already_knew == "none") %>% 
  mutate(acc_recall = as.numeric(acc_recall > 0),
         j_score = j_score - 0.7) %>% 
  lme4::glmer(acc_recall ~ j_score + (1 | subj_num), data = ., family = "binomial") %>%
  broom.mixed::tidy() %>%
  filter(effect == "fixed") %>% 
  select(term:statistic) %>% 
  mutate(across(where(is.numeric), signif, digits = 3)) %>% 
  knitr::kable()
```

]

.pull-right[

Formula: `acc_recall ~ j_score + interest + (1 + interest | subj_num)`

```{r}
retrieval %>% 
  filter(already_knew == "none") %>% 
  mutate(acc_recall = as.numeric(acc_recall > 0),
         # scaled to units of 0.1
         j_score = (j_score - 0.7) * 10,
         interest = (interest / 100) - .5) %>% 
  lme4::glmer(acc_recall ~ j_score + interest + (1 + interest | subj_num), data = ., family = "binomial") %>%
  broom.mixed::tidy() %>%
  filter(effect == "fixed") %>% 
  select(term:statistic) %>% 
  mutate(across(where(is.numeric), signif, digits = 3)) %>% 
  knitr::kable()
```

]

---

### Trivia experts _don't_ show better photo memory

```{r}
retrieval %>%
  # filter(already_knew == "none") %>% 
  mutate(across(c(resp_pic, resp_source),
                ~if_else(. > 0, 1L, 0L))) %>% 
  group_by(subj_num, j_score) %>% 
  summarize(acc_pic = mean(resp_pic),
            acc_source = mean(resp_source)) %>% 
  pivot_longer(cols = starts_with("acc"),
               names_to = "memory_type",
               values_to = "acc",
               names_prefix = "acc_") %>% 
  ggplot(aes(x = j_score, y = acc, color = memory_type, fill = memory_type)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(x = "Expertise score", y = " Mean accuracy WITHOUT confidence (0 to 1)") +
  theme_bw()
```

---

### _Photo_ recognition predicts _fact_ recall, slightly more so for experts

Quick and dirty maximum-likelihood logistic regression

Formula: `acc_recall ~ resp_pic * j_score + (1 + resp_pic | subj_num)`

```{r}
retrieval %>% 
  filter(already_knew == "none") %>% 
  mutate(acc_recall = as.numeric(acc_recall > 0),
         # across(c(resp_pic, resp_source), ~if_else(. > 0, 1L, 0L)),
         across(c(resp_pic, resp_source), ~(. / 100)),
         interest = (interest / 100) - 0.5,
         j_score = j_score - 0.7) %>% 
  lme4::glmer(acc_recall ~ resp_pic * j_score + interest + (1 + resp_pic | subj_num), data = ., family = "binomial") %>%
  broom.mixed::tidy() %>%
  filter(effect == "fixed") %>% 
  select(term:statistic) %>% 
  mutate(across(where(is.numeric), signif, digits = 3)) %>% 
  knitr::kable()
```

---

### _Source_ recognition predicts _fact_ recall for everyone

```{r}
retrieval %>% 
  filter(already_knew == "none") %>% 
  mutate(across(c(resp_pic, resp_source), ~if_else(. > 0, 1L, 0L)),
         acc_recall = if_else(acc_recall == 0.5, 1, acc_recall)) %>%
  group_by(subj_num, j_score, resp_source) %>%
  summarize(acc_recall = mean(acc_recall)) %>% 
  ggplot(aes(x = resp_source, y = acc_recall, color = fct_rev(cut_number(j_score, 3)), fill = fct_rev(cut_number(j_score, 3)))) + 
  geom_line(aes(group = subj_num), alpha = 0.1) +
  geom_smooth(method = "lm", alpha = 0.2) + 
  scale_x_continuous(breaks = 0:1) +
  scale_color_viridis_d(begin = 0.2, end = 0.9, direction = -1) +
  scale_fill_viridis_d(begin = 0.2, end = 0.9, direction = -1) +
  labs(x = "Source recognition accuracy",
       y = "Fact recall accuracy (generous)",
       color = "Expertise tertile",
       fill = "Expertise tertile") +
  theme_bw()
```

---

### _Source_ recognition predicts _fact_ recall for everyone

Quick and dirty maximum-likelihood logistic regression

Formula: `acc_recall ~ resp_source * j_score + (1 + resp_source | subj_num)`

```{r}
retrieval %>% 
  filter(already_knew == "none") %>% 
  mutate(acc_recall = as.numeric(acc_recall > 0),
         across(c(resp_pic, resp_source), ~if_else(. > 0, 1L, 0L)),
         j_score = j_score - 0.7) %>% 
  lme4::glmer(acc_recall ~ resp_source * j_score + (1 + resp_source | subj_num), data = ., family = "binomial") %>%
  broom.mixed::tidy() %>%
  filter(effect == "fixed") %>% 
  select(term:statistic) %>% 
  mutate(across(where(is.numeric), signif, digits = 3)) %>% 
  knitr::kable()
```

---

### _Combined_ recognition predicts _fact_ recall, slightly more so for experts

```{r}
retrieval %>% 
  filter(already_knew == "none") %>% 
  mutate(across(c(resp_pic, resp_source), ~if_else(. > 0, 1L, 0L)),
         resp_either = resp_pic + resp_source,
         acc_recall = as.integer(acc_recall > 0)) %>%
  filter(resp_either > 0) %>% 
  group_by(subj_num, j_score, resp_either) %>%
  summarize(acc_recall = mean(acc_recall)) %>% 
  ggplot(aes(x = resp_either, y = acc_recall, color = fct_rev(cut_number(j_score, 3)), fill = fct_rev(cut_number(j_score, 3)))) + 
  geom_line(aes(group = subj_num), alpha = 0.1) +
  geom_smooth(method = "lm", alpha = 0.2) + 
  scale_x_continuous(breaks = 1:2) +
  scale_color_viridis_d(begin = 0.2, end = 0.9, direction = -1) +
  scale_fill_viridis_d(begin = 0.2, end = 0.9, direction = -1) +
  labs(x = "Joint associate recognition accuracy",
       y = "Fact recall accuracy (generous)",
       color = "Expertise tertile",
       fill = "Expertise tertile") +
  theme_bw()
```

---

### _Combined_ recognition predicts _fact_ recall, slightly more so for experts

Quick and dirty maximum-likelihood logistic regression

Formula: `acc_recall ~ resp_either * j_score + (1 + resp_either | subj_num)`

```{r}
retrieval %>% 
  filter(already_knew == "none") %>% 
  mutate(acc_recall = as.numeric(acc_recall > 0),
         across(c(resp_pic, resp_source), ~if_else(. > 0, 1L, 0L)),
         # so, like 20% of participants have no retrieval trials where they missed on both source and pic
         # so this is run only using trials where people got _either_ source _or_ pic
         # aka the difference between getting one and getting both
         resp_either = resp_pic + resp_source - 1,
         j_score = j_score - 0.7) %>% 
  filter(resp_either > -1) %>% 
  lme4::glmer(acc_recall ~ resp_either * j_score + (1 + resp_either | subj_num), data = ., family = "binomial") %>%
  broom.mixed::tidy() %>%
  filter(effect == "fixed") %>% 
  select(term:statistic) %>% 
  mutate(across(where(is.numeric), signif, digits = 3)) %>% 
  knitr::kable()
```