### 2022-10-13 slides/results.Rmd
teeny updates before mariam meeting today

---
title: "Preliminary (quick and dirty) results"
subtitle: "J-Study"  
author: 
  - "Monica Thieu"
  - "Lauren Wilkins"
institute: "Columbia University Dept of Psychology"
date: 'Updated: `r Sys.Date()`'
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    nature:
      slideNumberFormat: "%current%"
      highlightStyle: github
      highlightLines: true
      ratio: 16:9
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(
  dev = "svglite", fig.width=9, fig.height=3.5, fig.retina=3,
  out.width = "100%",
  cache = FALSE,
  echo = FALSE,
  message = FALSE, 
  warning = FALSE,
  hiline = TRUE
)

require(xaringanthemer)
source(here::here("R", "get_cleaned_data.R"))

count_streak <- function (acc_col, min_streak_length = 3) {
  streak_count <- 0
  current_streak <- 0
  for (i in 1:length(acc_col)) {
    if (acc_col[i] == 0) current_streak <- 0
    if (acc_col[i] == 1) {
      current_streak <- current_streak + acc_col[i]
      if (current_streak == min_streak_length) {
        # Only triggered once per streak because using == not >=
        streak_count <- streak_count + 1
      }
    }
  }
  return (streak_count)
}
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
style_mono_accent(
  base_color = "#0046a6",
  text_font_google = google_font("Lora"),
  header_font_google = google_font("Nunito"),
  code_font_google = google_font("Fira Code")
)
```

class: inverse, middle

# Proposed main figures

---

### Trivia experts _don't_ show better photo memory

```{r}
retrieval %>%
  filter(already_knew == "none") %>% 
  mutate(across(c(resp_pic, resp_source, acc_recall),
                ~if_else(. > 0, 1L, 0L))) %>% 
  group_by(subj_num, j_score) %>% 
  summarize(acc_pic = mean(resp_pic),
            acc_source = mean(resp_source),
            acc_fact = mean(acc_recall)) %>% 
  pivot_longer(cols = starts_with("acc"),
               names_to = "memory_type",
               values_to = "acc",
               names_prefix = "acc_") %>% 
  ggplot(aes(x = j_score, y = acc, color = memory_type, fill = memory_type)) +
  geom_hline(yintercept = 0.5, linetype = "dotted") +
  geom_point() +
  geom_smooth(method = "lm") +
  facet_wrap(~ memory_type) +
  guides(color = "none", fill = "none") +
  labs(x = "Expertise score", y = " Mean accuracy WITHOUT confidence (0 to 1)",
       title = "Memory as a function of trivia expertise",
       subtitle = "Only trials with fully novel facts") +
  theme_bw()
```

---

### _Photo_ recognition predicts _fact_ recall, slightly more so for experts

```{r}
retrieval %>% 
  filter(already_knew == "none") %>% 
  mutate(across(c(resp_pic, resp_source), ~if_else(. > 0, 1L, 0L)),
         acc_recall = if_else(acc_recall == 0.5, 1, acc_recall)) %>%
  group_by(subj_num, j_score, resp_pic) %>%
  summarize(acc_recall = mean(acc_recall)) %>% 
  ggplot(aes(x = resp_pic, y = acc_recall, color = fct_rev(cut_number(j_score, 2)), fill = fct_rev(cut_number(j_score, 1)))) + 
  geom_line(aes(group = subj_num), alpha = 0.1) +
  geom_smooth(method = "lm", alpha = 0.2) + 
  scale_x_continuous(breaks = 0:1) +
  scale_color_viridis_d(begin = 0.2, end = 0.9, direction = -1) +
  scale_fill_viridis_d(begin = 0.2, end = 0.9, direction = -1) +
  labs(x = "Photo recognition accuracy",
       y = "Fact recall accuracy (generous)",
       color = "Expertise median split",
       fill = "Expertise median split") +
  theme_bw()
```
???

Main figure subplots: coefplot, a better version of this with median split and as close to trialwise data as possible, with Bayesian model fit ribbons on top

---

### _Source_ recognition predicts _fact_ recall for everyone

```{r}
retrieval %>% 
  filter(already_knew == "none") %>% 
  mutate(across(c(resp_pic, resp_source), ~if_else(. > 0, 1L, 0L)),
         acc_recall = if_else(acc_recall == 0.5, 1, acc_recall)) %>%
  group_by(subj_num, j_score, resp_source) %>%
  summarize(acc_recall = mean(acc_recall)) %>% 
  ggplot(aes(x = resp_source, y = acc_recall, color = fct_rev(cut_number(j_score, 3)), fill = fct_rev(cut_number(j_score, 3)))) + 
  geom_line(aes(group = subj_num), alpha = 0.1) +
  geom_smooth(method = "lm", alpha = 0.2) + 
  scale_x_continuous(breaks = 0:1) +
  scale_color_viridis_d(begin = 0.2, end = 0.9, direction = -1) +
  scale_fill_viridis_d(begin = 0.2, end = 0.9, direction = -1) +
  labs(x = "Source recognition accuracy",
       y = "Fact recall accuracy (generous)",
       color = "Expertise tertile",
       fill = "Expertise tertile") +
  theme_bw()
```

???

Same plot subfigures as the previous

Also, see about running a model with pic/source as a covariate to explicitly test the difference in interactions

---

class: inverse, middle

# Demographics

---

### More men than women, but age does not appear correlated with gender

```{r}
q_google_demos %>% 
  left_join(q_google_demos %>%
              count(gender) %>%
              mutate(gender_n = glue::glue("{gender} (n = {n})")) %>%
              select(-n),
            by = "gender") %>% 
  ggplot(aes(x = age, fill = gender_n)) +
  geom_histogram(position = "identity", alpha = 0.5, binwidth = 2) +
  scale_y_continuous(breaks = scales::breaks_extended(6))
```

---

### What is our sample spread of trivia expertise?

```{r}
j_exp %>% 
  ggplot(aes(x = j_score)) +
  geom_histogram(binwidth = .05) +
  geom_vline(xintercept = 0.7, linetype = "dotted") +
  annotate("text",
           x = .67,
           y = 20,
           label = "Approx cutoff for Jeopardy callback",
           hjust = 1) +
  labs(x = "Expertise score (fraction correct out of 50 questions)",
       y = "count")
```

---

class: inverse, middle

# Encoding

---

### Trivia experts tend to know more encoding facts already

```{r}
encoding %>% 
  group_by(subj_num, j_score, group) %>% 
  summarize(pct_novel = sum(!is.na(already_knew) & already_knew == "none")) %>% 
  ggplot(aes(x = j_score, y = pct_novel, color = group, fill = group)) +
  geom_hline(yintercept = 40, linetype = "dotted") +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(x = "Trivia expertise", y = "# museum facts rated as novel") +
  theme_bw()
```

---

### Trivia experts might be more interested at encoding, but not by much

```{r}
this_cor <- encoding %>%
  group_by(subj_num, j_score) %>% 
  summarize(interest = median(resp)) %$% 
  cor(j_score, interest) %>% 
  round(2)
```

```{r}
encoding %>% 
  group_by(subj_num, j_score) %>% 
  summarize(resp_median = median(resp),
            q25 = quantile(resp, .25),
            q75 = quantile(resp, .75)) %>% 
  ggplot(aes(x = j_score, y = resp_median)) +
  geom_errorbar(aes(ymin = q25, ymax = q75), width = 0, alpha = 0.5) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(x = "Expertise score",
       y = "P's median interest (across facts)") +
  theme_bw()
```

---

### Trivia experts might be more interested at encoding, but not by much

Quick and dirty maximum-likelihood linear regression

Formula: `interest ~ j_score + already_knew + (1 | subj_num)`

```{r}
encoding %>% 
  mutate(j_score = (j_score - 0.7) * 10) %>% 
  lme4::lmer(resp ~ j_score + already_knew + (1 | subj_num), data = .) %>% 
  broom.mixed::tidy() %>%
  filter(effect == "fixed") %>% 
  select(term:statistic) %>% 
  mutate(across(where(is.numeric), signif, digits = 3)) %>% 
  knitr::kable()
```

---

class: inverse, middle

# Retrieval

---

### Recall is higher when people already know the facts (duh)

Accordingly, all subsequent results involving fact retrieval are shown only for facts rated as novel at encoding.

Excluding facts known at encoding should not introduce collider bias into the retrieval results, because recall cannot cause prior knowledge at encoding.

```{r}
retrieval %>% 
  filter(!is.na(already_knew)) %>% 
  group_by(subj_num, already_knew) %>% 
  summarize(acc_recall = mean(acc_recall)) %>% 
  ggplot(aes(x = already_knew, y = acc_recall)) +
  geom_violin() +
  labs(x = "Fact prior knowledge at encoding", y = "% correct recall") +
  theme_bw()
```

---

### Trivia experts show better recall for novel facts

```{r}
retrieval %>% 
  filter(already_knew == "none") %>% 
  group_by(subj_num, j_score) %>% 
  summarize(acc_recall = mean(acc_recall)) %>% 
  ggplot(aes(x = j_score, y = acc_recall)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(x = "Expertise score", y = "% correct recall for novel facts") +
  theme_bw()
```

---

### Trivia experts' fact recall is _not_ driven by interest at encoding

.pull-left[

Formula: `acc_recall ~ j_score + (1 | subj_num)`

```{r}
retrieval %>% 
  filter(already_knew == "none") %>% 
  mutate(acc_recall = as.numeric(acc_recall > 0),
         j_score = (j_score - 0.7) * 10) %>% 
  lme4::glmer(acc_recall ~ j_score + (1 | subj_num), data = ., family = "binomial") %>%
  broom.mixed::tidy() %>%
  filter(effect == "fixed") %>% 
  select(term:statistic) %>% 
  mutate(across(where(is.numeric), signif, digits = 3)) %>% 
  knitr::kable()
```

]

.pull-right[

Formula: `acc_recall ~ j_score + interest + (1 + interest | subj_num)`

```{r}
retrieval %>% 
  filter(already_knew == "none") %>% 
  mutate(acc_recall = as.numeric(acc_recall > 0),
         # scaled to units of 0.1
         j_score = (j_score - 0.7) * 10,
         # scaled to units of 0.1 where 1 is the full scale
         interest = (interest / 10) - 5) %>% 
  lme4::glmer(acc_recall ~ j_score + interest + (1 + interest | subj_num), data = ., family = "binomial") %>%
  broom.mixed::tidy() %>%
  filter(effect == "fixed") %>% 
  select(term:statistic) %>% 
  mutate(across(where(is.numeric), signif, digits = 3)) %>% 
  knitr::kable()
```

]

---

### _Photo_ recognition predicts _fact_ recall, slightly more so for experts

Quick and dirty maximum-likelihood logistic regression

Formula: `acc_recall ~ resp_pic * j_score + (1 + resp_pic | subj_num)`

```{r}
retrieval %>% 
  filter(already_knew == "none") %>% 
  mutate(acc_recall = as.numeric(acc_recall > 0),
         # across(c(resp_pic, resp_source), ~if_else(. > 0, 1L, 0L)),
         across(c(resp_pic, resp_source), ~(. / 100)),
         interest = (interest / 10) - 5,
         j_score = (j_score - 0.7) * 10) %>% 
  lme4::glmer(acc_recall ~ resp_pic * j_score + interest + (1 + resp_pic | subj_num), data = ., family = "binomial") %>%
  broom.mixed::tidy() %>%
  filter(effect == "fixed") %>% 
  select(term:statistic) %>% 
  mutate(across(where(is.numeric), signif, digits = 3)) %>% 
  knitr::kable()
```

---

### _Source_ recognition predicts _fact_ recall for everyone

Quick and dirty maximum-likelihood logistic regression

Formula: `acc_recall ~ resp_source * j_score + (1 + resp_source | subj_num)`

```{r}
retrieval %>% 
  filter(already_knew == "none") %>% 
  mutate(acc_recall = as.numeric(acc_recall > 0),
         across(c(resp_pic, resp_source), ~if_else(. > 0, 1L, 0L)),
         j_score = (j_score - 0.7) * 10) %>% 
  lme4::glmer(acc_recall ~ resp_source * j_score + (1 + resp_source | subj_num), data = ., family = "binomial") %>%
  broom.mixed::tidy() %>%
  filter(effect == "fixed") %>% 
  select(term:statistic) %>% 
  mutate(across(where(is.numeric), signif, digits = 3)) %>% 
  knitr::kable()
```

---

### _Combined_ recognition predicts _fact_ recall, slightly more so for experts

```{r}
retrieval %>% 
  filter(already_knew == "none") %>% 
  mutate(across(c(resp_pic, resp_source), ~if_else(. > 0, 1L, 0L)),
         resp_either = resp_pic + resp_source,
         acc_recall = as.integer(acc_recall > 0)) %>%
  filter(resp_either > 0) %>% 
  group_by(subj_num, j_score, resp_either) %>%
  summarize(acc_recall = mean(acc_recall)) %>% 
  ggplot(aes(x = resp_either, y = acc_recall, color = fct_rev(cut_number(j_score, 3)), fill = fct_rev(cut_number(j_score, 3)))) + 
  geom_line(aes(group = subj_num), alpha = 0.1) +
  geom_smooth(method = "lm", alpha = 0.2) + 
  scale_x_continuous(breaks = 1:2) +
  scale_color_viridis_d(begin = 0.2, end = 0.9, direction = -1) +
  scale_fill_viridis_d(begin = 0.2, end = 0.9, direction = -1) +
  labs(x = "Joint associate recognition accuracy",
       y = "Fact recall accuracy (generous)",
       color = "Expertise tertile",
       fill = "Expertise tertile") +
  theme_bw()
```

---

### _Combined_ recognition predicts _fact_ recall, slightly more so for experts

Quick and dirty maximum-likelihood logistic regression

Formula: `acc_recall ~ resp_either * j_score + (1 + resp_either | subj_num)`

```{r}
retrieval %>% 
  filter(already_knew == "none") %>% 
  mutate(acc_recall = as.numeric(acc_recall > 0),
         across(c(resp_pic, resp_source), ~if_else(. > 0, 1L, 0L)),
         # so, like 20% of participants have no retrieval trials where they missed on both source and pic
         # so this is run only using trials where people got _either_ source _or_ pic
         # aka the difference between getting one and getting both
         resp_either = resp_pic + resp_source - 1,
         j_score = (j_score - 0.7) * 10) %>% 
  filter(resp_either > -1) %>% 
  lme4::glmer(acc_recall ~ resp_either * j_score + (1 + resp_either | subj_num), data = ., family = "binomial") %>%
  broom.mixed::tidy() %>%
  filter(effect == "fixed") %>% 
  select(term:statistic) %>% 
  mutate(across(where(is.numeric), signif, digits = 3)) %>% 
  knitr::kable()
```