### 2024-01-27 writing/supplement.Rmd
updates per pbr reviewers

---
title: "Supplementary material"
subtitle: "Main paper: Episodic memory bolsters acquisition of new semantic knowledge in trivia experts"
author: "MT, LW, & MA"
date: "Updated `r Sys.Date()`"
output:
  pdf_document:
    toc: true
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
require(targets)
require(tidyverse)
require(cowplot)
require(magrittr)
require(rlang)
source(here::here("R", "analyze", "make_plots_ms.R")) # for recode_params_pretty

target_store <- here::here("ignore", "_targets")
signif_digits <- 3L

summarize_kable <- function (data,
                             col,
                             digs = signif_digits,
                             fn_list = list(mean = \(x) mean(x, na.rm = T),
                                            sd = \(x) sd(x, na.rm = T))) {
  data %>% 
    summarize(across({{col}}, fn_list, .names = "{.fn}")) %>% 
    mutate(across(where(is.numeric), \(x) signif(x, digits = digs))) %>% 
    knitr::kable()
}

fn_list_model.coefs <- list(`95% CI lower bound` = \(x) quantile(x, .025),
                            Median = \(x) median(x, na.rm = T),
                            `95% CI upper bound` = \(x) quantile(x, .975))

# Numerical stats mentioned in the main text are also tabulated in this document
# They are in the chunks tagged with include=FALSE so that they don't appear in the supplement
# Those chunks are only intended to be run interactively, and then (sigh)
# the outputs are to be copied and pasted into the main text... it was the easiest overall
```

## Demographics

```{r include=FALSE}
# Numbers for the main text
q_google_demos <- tar_read(q_google_demos, store = target_store)
j_exp <- tar_read(expertise_summarized, store = target_store)

nrow(q_google_demos)
```

### Figure 1: Participant demographics

```{r}
knitr::include_graphics(here::here("ignore", "figs", "ms_fig3.png"))
```

Participant demographics. **(A)** 132 total participants aged 18 and up enrolled in the study (M~age~ = `r signif(mean(q_google_demos$age, na.rm = TRUE), digits = 3)` yrs, SD~age~ = `r signif(sd(q_google_demos$age, na.rm = TRUE), digits = 3)` yrs). Histogram bars are stacked by participant gender such that the total bar height reflects all participants in that age bin. **(B)** Participants ranged in pre-existing trivia expertise, as measured by their score out of 50 questions on a cued-recall general knowledge test. The median expertise score was 36/50 questions correct.

### Figure 2: Trivia expertise as a function of gender

```{r, include=FALSE}
# These stats are in the paper!
q_google_demos %>% 
  count(gender) %>% 
  knitr::kable()
```

```{r}
q_google_demos %>% 
  left_join(j_exp %>% mutate(subj_num = factor(subj_num))) %>% 
  mutate(gender = coalesce(gender, "Not reported")) %>% 
  ggplot(aes(x = j_score, fill = gender)) +
  geom_histogram(position = "identity", alpha = 0.5) +
  labs(x = "Trivia expertise score") +
  theme_bw()
```

Men in our sample tended to score higher than women and genderqueer participants. LearnedLeague, the website from which we recruited, has more men than women participants, so we do not interpret these results any further.

### Table 1: Participant breakdown by race

```{r}
# These stats are in the paper!
q_google_demos %>% 
  # Anonymizing this one friend
  mutate(race = if_else(race == "Asian, I want to be clear that I am Indian and not East Asian", "Asian", race)) %>% 
  count(race) %>% 
  arrange(desc(n)) %>% 
  knitr::kable()
```

The vast majority of participants reported their race as White, with the largest minority of participants reporting their race as Asian/Pacific Islander. While LearnedLeague does not collect race information on members, we expect that this distribution is representative of website membership. We do not use race in any subsequent analyses because of the low race diversity in the sample.

```{r age-mean-sd, include=FALSE}
# These stats are in the paper!
q_google_demos %>% 
  summarize_kable(age)
```

```{r plot-jscore-by-age, include=FALSE}
# TBD whether this should be rendered as a supplemental figure
q_google_demos %>% 
  inner_join(j_exp %>% 
               mutate(subj_num = factor(subj_num)), 
             by = "subj_num") %>% 
  lm(scale(j_score) ~ scale(age), data = .) %>% 
  summary()

q_google_demos %>% 
  inner_join(j_exp %>% 
               mutate(subj_num = factor(subj_num)), 
             by = "subj_num") %>% 
  ggplot(aes(x = age, y = j_score)) + 
  geom_point() + 
  geom_smooth(method = "lm")
```

\newpage

## Expertise pre-test

After completing the general knowledge pre-test, we asked participants to rate their meta-memory for each question. Ratings were based on Pereverseff and Bodner (2020)'s ratings. Participants were instructed as follows:

- "Learning memory": When the answer came to mind, you also remembered a specific prior experience of learning this fact (e.g., how, when, and/or where you learned it). It does not have to be a memory of the first time you learned it.
- "Related memory": When the answer came to mind, you also remembered a specific related personal memory, but the memory was not about learning this fact.
- "Just know": When the answer came to mind, a specific personal memory did not come to mind. Instead, the answer just seems to be part of your general knowledge.
- "Guess": You do not experience a specific personal memory or a general feeling of knowing the answer. It is an educated guess that you may have arrived at with some reasoning.
- "Don't know": You have no idea what the answer could be, and you cannot even make a guess. You do not experience a specific personal memory or a general feeling of knowing the answer.

After participants gave their metamemory ratings, we randomly selected one general knowledge fact endorsed by the participant for each of the five potential labels,
and asked them to describe their metamemory state for that fact in words. We did not analyze any of the metamemory ratings because our validity checks indicated that participants did not use the ratings as instructed. Specifically, when we inspected participants’ descriptions, we found enough responses indicating “just know”-style metamemory for memories that participants had tagged with the “remember” responses (“learning memory” and “related memory”) that we considered the data to be unusable based on participants failing to follow task instructions.

\newpage

## Encoding task

```{r include=FALSE}
# These stats are in response to PB&R Reviewer 2's question about the encoding categories
encoding <- tar_read(encoding, store = target_store)

encoding_categories <- encoding %>% 
  distinct(subj_num, j_score, group, category) %>% 
  mutate(category = fct_relevel(category,
                                "arms", "gems", "musi"),
         category = fct_recode(category,
                               "Arms and armor" = "arms",
                               "Gemstone geology" = "gems",
                               "Historical musical instruments" = "musi",
                               "Prehistoric animals" = "dino",
                               "How cars work" = "cars",
                               "Cooking tools and techniques" = "cook"))

encoding_categories %>% 
  count(group, category) %>% 
  group_by(group) %>% 
  mutate(prob = n/sum(n))
```

```{r include=FALSE}
# These stats are in the paper!
encoding %>%
  # there is literally _one_ trial where the timing seems to have messed up and run over
  filter(rt <= 35100) %>% 
  mutate(rt = rt / 1000) %>% 
  summarize_kable(rt)
```

```{r include=FALSE}
# These stats are in the paper!
j_exp %>% 
  # to get it in units of questions
  mutate(j_score = j_score * 50) %>% 
  summarize_kable(j_score,
                  fn_list = list(min = \(x) min(x, na.rm = T),
                                 max = \(x) max(x, na.rm = T),
                                 median = \(x) median(x, na.rm = T)))
```

```{r include=FALSE}
# These stats are in the paper!
encoding %>% 
  filter(already_knew == "none") %>% 
  count(subj_num) %>% 
  summarize_kable(n,
                  digs = 2L,
                  fn_list = list(min = \(x) min(x, na.rm = T),
                                 mean = \(x) mean(x, na.rm = T),
                                 sd = \(x) sd(x, na.rm = T)))
```

### Table 2: Model coefficients for encoding familiarity as a function of trivia expertise

```{r}
tar_read(preplot_params_alreadyknew, store = target_store) %>%
  mutate(term = recode_params_pretty(term)) %>% 
  group_by(term) %>% 
  summarize_kable(estimate,
                  fn_list = fn_list_model.coefs)
```

This model was run as a Bayesian multilevel logistic regression (with the same model setting as the main model reported in the main text) predicting fact familiarity for each fact (unfamiliar = 0, familiar = 1) as a function of trivia expertise, with a random intercept for each participant. Participants rated their familiarity as "none", "some", or "all" for each encoding fact. For this analysis, we binned "some" and "all" together to compare the totally unfamiliar facts against facts for which participants reported _at least_ some familiarity.

### Table 3: Model coefficients for encoding interest as a function of trivia expertise

```{r}
tar_read(preplot_params_interest, store = target_store) %>%
  mutate(term = recode_params_pretty(term),
         term = fct_relevel(term, "Intercept")) %>% 
  group_by(term) %>% 
  summarize_kable(estimate,
                  fn_list = fn_list_model.coefs)
```

This model was run as a Bayesian multilevel _linear_ regression (with the same model settings as the main model reported in the main text) predicting interest at encoding (on a 0-100 point scale, centered at 50 pts) as a function of trivia expertise, with a random intercept for each participant. We also included a covariate for first vs. second museum, coded the same way as it was for the model in the main text.

For this model, effective N failed to reach 10% of the sampling iterations for several parameters, even after increasing the number of sampling iterations to 3000 iterations/chain. However, no other associated model reliability parameters (like Rhat) suggested concern. We thus report these model outcomes as is. We expect that some of the apparently low N~eff~ may arise from individual participants having different distributions of interest ratings across their facts.

\newpage

## Retrieval tasks

```{r load-retrieval}
retrieval <- tar_read(retrieval, store = target_store) %>% 
  left_join(q_google_demos, by = "subj_num")
```

### Figure 3: Recall as a function of semantic category

```{r plot-recall-by-category}
knitr::include_graphics(here::here("ignore", "figs", "ms_fig_supp_encoding_categories.png"))
```

**(A)** Median trivia expertise of participants assigned to each category is roughly evenly distributed around the whole-sample median trivia expertise (marked on the graph with a reference line). The number of participants assigned to each category is also visible (number of dots for each category). Across participants, the three “academic” categories were assigned roughly evenly, while among the three “non-academic” categories, prehistoric animal facts were presented to about 50% of participants. **(B)** Each dot represents one participant’s recall accuracy for novel facts, with each participant having one dot per color for their “academic” and “nonacademic” categories. Within the “academic” topics, recall performance was not significantly different across categories. Within the “non-academic” topics, performance in the car parts/functions category tended to be superior to that in the prehistoric animals and cooking tools/techniques categories (see following two Tables).

\newpage

### Table 4: Model coefficients for recall as a function of (academic) category and expertise 
```{r}
# These stats are in response to PB&R reviewers
tar_read(preplot_params_fact_academic, store = target_store) %>%
  mutate(term = recode_params_pretty(term),
         term = fct_relevel(term, "Intercept"),
         term = fct_recode(term,
                           "Arms vs. gems" = "category_arms",
                           "Musical instruments vs. gems" = "category_musi")) %>% 
  group_by(term) %>% 
  summarize_kable(estimate,
                  fn_list = fn_list_model.coefs)
```

"Academic" fact categories all showed similar recall performance. This model was run as a Bayesian multilevel logistic regression (with the same model settings as the main model reported in the main text) predicting novel "academic" fact recall (half of trials) as a function of fact category and trivia expertise, with a random intercept for each participant. Gemstone geology was coded as the baseline category, so coefficients reflect differences between gems vs. arms & armor, and gems vs. historical musical instruments. We also included covariates for interest at encoding and first vs. second museum, coded the same way as they were for the model in the main text.

### Table 5: Model coefficients for recall as a function of (nonacademic) category and expertise

```{r}
# These stats are in response to PB&R reviewers
tar_read(preplot_params_fact_nonacademic, store = target_store) %>%
  mutate(term = recode_params_pretty(term),
         term = fct_relevel(term, "Intercept"),
         term = fct_recode(term,
                           "Cooking vs. cars" = "category_cook",
                           "Prehistoric animals vs. cars" = "category_dino")) %>% 
  group_by(term) %>% 
  summarize_kable(estimate,
                  fn_list = fn_list_model.coefs)
```

Of "nonacademic" fact categories, recall performance was higher for car facts, but independently of trivia expertise. This model was run as a Bayesian multilevel logistic regression (with the same model settings as the main model reported in the main text) predicting novel "nonacademic" fact recall (half of trials) as a function of fact category and trivia expertise, with a random intercept for each participant. Car parts/function was coded as the baseline category, so coefficients reflect differences between cars vs. cooking techniques, and cars vs. prehistoric animals. We also included covariates for interest at encoding and first vs. second museum, coded the same way as they were for the model in the main text.

```{r recall-rt-start-by-age, include=FALSE}
# These stats are in response to PB&R reviewers
retrieval %>% 
  lme4::lmer(rt_start_recall ~ scale(age, scale=F) + (1 | subj_num), data = .) %>% 
  summary()
```

```{r recall-rt-end-by-age, include=FALSE}
# These stats are in response to PB&R reviewers
retrieval %>% 
  lme4::lmer(rt_end_recall ~ scale(age, scale=F) + (1 | subj_num), data = .) %>% 
  summary()
```

```{r recall-rt-timeout-by-age, include=FALSE}
# These stats are in response to PB&R reviewers
retrieval %>% 
  filter(rt_end_recall >= 15000) %>% 
  count(subj_num, age) %>% 
  lm(n ~ scale(age, scale=F), data = .) %>% 
  summary()
```

```{r encoding-retrieval-gap, include=FALSE}
retrieval %>% 
  select(subj_num, encoding_trial_num, retrieval_trial_num) %>% 
  # estimated encoding trial length: 27 s (counting already knew and fixation)
  mutate(encoding_trial_time = (encoding_trial_num - 1) * 27, 
         # estimated retrieval trial length: 8 s, plus 5 min wait plus encoding offset
         retrieval_trial_time = (retrieval_trial_num - 1) * 8 + 300 + max(encoding_trial_time), 
         delay = retrieval_trial_time - encoding_trial_time) %>% 
  summarize(mean_delay_min = mean(delay/60), sd_delay_min = sd(delay/60)) %>% 
  mutate(across(where(is.numeric), \(x) signif(x, digits = 3)))
```

```{r preplot-params-fact-by-both-coefs, include=FALSE}
# These stats are in the paper!
preplot_params_fact_by_both <- tar_read(preplot_params_fact_by_both, store = target_store)

preplot_params_fact_by_both %>%
  group_by(term) %>% 
  summarize_kable(estimate,
                  fn_list = fn_list_model.coefs)
```

```{r preplot-params-fact-by-both-simple-fx, include=FALSE}
# These stats are in the paper!
# "low" trivia expertise for the preplots is j_score = -2
# while "high" trivia expertise is +1
preplot_params_fact_by_both %>% 
  filter(term %in% c("resp_pic:resp_source", "resp_pic:resp_source:j_score")) %>% 
  mutate(term = str_replace_all(term, ":", ".")) %>% 
  pivot_wider(names_from = term,
              values_from = estimate) %>% 
  mutate(low = resp_pic.resp_source - 2*resp_pic.resp_source.j_score,
         high = resp_pic.resp_source + resp_pic.resp_source.j_score) %>% 
  select(iteration, low, high) %>% 
  pivot_longer(cols = c(low, high),
               names_to = "j_score",
               values_to = "estimate") %>% 
  group_by(j_score) %>% 
  summarize_kable(estimate,
                  fn_list = fn_list_model.coefs)
```

```{r include=FALSE}
# These stats are in the paper!
tar_read(preplot_fixef_fact_by_both, store = target_store) %>% 
  group_by(j_score, resp_pic, resp_source) %>% 
  summarize_kable(acc_pred,
                  fn_list = list(median = median))
```

```{r recall-rt-jscore, include=FALSE}
# These stats are in the response to PB&R reviewers
retrieval %>% 
  # filter(acc_recall > 0) %>% 
  mutate(across(starts_with("resp"), \(x) as.integer(x > 0)),
         j_score = (j_score - 0.7) * 10) %>% 
  lme4::glmer(rt_start_recall ~ j_score + (1 | subj_num), data = .) %>% 
  broom.mixed::tidy()
```

```{r resp-pic-recall-rt, include=FALSE}
# These stats are in the response to PB&R reviewers
retrieval %>% 
  filter(acc_recall > 0) %>% 
  mutate(across(starts_with("resp"), \(x) as.integer(x > 0)),
         j_score = (j_score - 0.7) * 10) %>% 
  lme4::glmer(resp_pic ~ scale(rt_start_recall) * j_score + (1 | subj_num),
              data = ., family = binomial) %>% 
  broom.mixed::tidy()
```

```{r preplot-params-resp-pic-recall-rt-coefs, include=FALSE}
# These stats are in the paper! After request by PB&R reviewers
preplot_params_pic_rt.recall <- tar_read(preplot_params_pic_rt.recall, store = target_store)

preplot_params_pic_rt.recall %>%
  group_by(term) %>% 
  summarize_kable(estimate,
                  fn_list = fn_list_model.coefs)
```

```{r preplot-params-resp-pic-recall-rt-simple-fx, include=FALSE}
# These stats are in the paper! After request by PB&R reviewers
# "low" trivia expertise for the preplots is j_score = -2
# while "high" trivia expertise is +1
preplot_params_pic_rt.recall %>% 
  filter(term %in% c("rt_start_recall", "rt_start_recall:j_score")) %>% 
  mutate(term = str_replace_all(term, ":", ".")) %>% 
  pivot_wider(names_from = term,
              values_from = estimate) %>% 
  mutate(low = rt_start_recall - 2*rt_start_recall.j_score,
         high = rt_start_recall + rt_start_recall.j_score) %>% 
  select(iteration, low, high) %>% 
  pivot_longer(cols = c(low, high),
               names_to = "j_score",
               values_to = "estimate") %>% 
  group_by(j_score) %>% 
  summarize_kable(estimate,
                  fn_list = fn_list_model.coefs)
```

```{r preplot-params-resp-source-recall-rt-coefs, include=FALSE}
# These stats are in the paper! After request by PB&R reviewers
preplot_params_source_rt.recall <- tar_read(preplot_params_source_rt.recall, store = target_store)

preplot_params_source_rt.recall %>%
  group_by(term) %>% 
  summarize_kable(estimate,
                  fn_list = fn_list_model.coefs)
```

```{r preplot-params-resp-source-recall-rt-simple-fx, include=FALSE}
# These stats are in the paper! After request by PB&R reviewers
# "low" trivia expertise for the preplots is j_score = -2
# while "high" trivia expertise is +1
preplot_params_source_rt.recall %>% 
  filter(term %in% c("rt_start_recall", "rt_start_recall:j_score")) %>% 
  mutate(term = str_replace_all(term, ":", ".")) %>% 
  pivot_wider(names_from = term,
              values_from = estimate) %>% 
  mutate(low = rt_start_recall - 2*rt_start_recall.j_score,
         high = rt_start_recall + rt_start_recall.j_score) %>% 
  select(iteration, low, high) %>% 
  pivot_longer(cols = c(low, high),
               names_to = "j_score",
               values_to = "estimate") %>% 
  group_by(j_score) %>% 
  summarize_kable(estimate,
                  fn_list = fn_list_model.coefs)
```

```{r subset-correct-recall-details, include=FALSE}
# These stats are also in the response to PB&R reviewers
retrieval %>% 
  filter(acc_recall > 0) %>% 
  mutate(across(starts_with("resp"), \(x) as.integer(x > 0)),
         no_details = as.integer(resp_source == 0 & resp_pic == 0),
         j_score = (j_score - 0.7) * 10) %>% 
  lme4::glmer(no_details ~ j_score + (1 | subj_num),
              data = .,
              family = binomial) %>% 
  summary()
```

### Table 6: Model coefficients for photo memory as a function of trivia expertise

```{r}
tar_read(preplot_params_pic, store = target_store) %>%
  mutate(term = recode_params_pretty(term)) %>% 
  group_by(term) %>% 
  summarize_kable(estimate,
                  fn_list = fn_list_model.coefs)
```

This model was run as a Bayesian multilevel logistic regression (with the same model settings as the main model reported in the main text) predicting binary photo recognition memory as a function of trivia expertise, with a random intercept for each participant. We also included covariates for interest at encoding and first vs. second museum, coded the same way as they were for the model in the main text.

### Table 7: Model coefficients for museum memory as a function of trivia expertise

```{r}
tar_read(preplot_params_source, store = target_store) %>%
  mutate(term = recode_params_pretty(term)) %>% 
  group_by(term) %>% 
  summarize_kable(estimate,
                  fn_list = fn_list_model.coefs)
```

This model was run as a Bayesian multilevel logistic regression (with the same model settings as the main model reported in the main text) predicting binary museum recognition memory as a function of trivia expertise, with a random intercept for each participant. We also included covariates for interest at encoding and first vs. second museum, coded the same way as they were for the model in the main text.

### Table 8: Model coefficients for fact recall as a function of expertise, for _familiar_ facts

```{r}
# Values from this table are also reported in the response to PB&R reviewers
preplot_params_fact_by_both_known.facts <- tar_read(preplot_params_fact_by_both_known.facts, store = target_store)

preplot_params_fact_by_both_known.facts %>% 
  mutate(term = recode_params_pretty(term),
         term = fct_relevel(term,
                              "Intercept",
                              "Interest at encoding",
                              "Trivia expertise",
                              "First vs. second museum",
                              "Photo memory",
                              "Museum memory",
                              "Expertise x photo memory",
                              "Expertise x museum memory",
                              "Photo memory x museum memory",
                              "Photo x museum x expertise")) %>% 
  group_by(term) %>% 
    summarize_kable(estimate,
                    fn_list = fn_list_model.coefs)
```

This model was run as a Bayesian multilevel logistic regression, with the same model settings as the main model reported in the main text. This model has the exact same outcome and predictors as the main model, but has been fit only to trials where participants reported knowing "some" or "all" of the fact prior to the current study. Critically, we do not find the same three-way interaction effect of photo memory, museum memory, and trivia expertise when analyzing recall performance only for previously familiar facts.

```{r include=FALSE}
# These simple effect stats are in the response to PB&R reviewers
preplot_params_fact_by_both_known.facts %>% 
  filter(term %in% c("resp_pic", "resp_source", "resp_pic:j_score", "resp_source:j_score")) %>% 
  mutate(term = str_replace_all(term, ":", ".")) %>% 
  separate_wider_delim(term, delim = ".", names = c("term", "coef_type"), too_few = "align_start") %>% 
  mutate(coef_type = if_else(is.na(coef_type), "main", "intxn")) %>% 
  pivot_wider(names_from = coef_type,
              values_from = estimate) %>% 
  mutate(low = main - 2*intxn,
         high = main + intxn) %>% 
  select(iteration, term, low, high) %>% 
  pivot_longer(cols = c(low, high),
               names_to = "j_score",
               values_to = "estimate") %>% 
  group_by(term, j_score) %>% 
  summarize_kable(estimate,
                  fn_list = fn_list_model.coefs)
```